---
title: "HH_number sense check and data change"
author: "Mark Ruddy"
date: "17 July 2016"
output: html_document
---

## HH_number check

Instruction: Sense check ? >8 is odd. Cross check with other person number data.

### Check values in HH_number variable.
Anything >8 needs investigating.
```{r Value HH_number}
ppv %>%
  filter(HH_number>8, is.na(HH_text)) %>%
  select(ID) %>%
  View()
## The following IDs don't have good explanations for >8 household members: 1257, 1635, 1785, 2441, 3213, 5090

## Missing value=-99
ppv %>%
  select(HH_number) %>%
  distinct() %>%
  arrange(HH_number)
## There are no missing values coded as -99.

## NAs
ppv %>%
  select(HH_number) %>%
  filter(is.na(HH_number)) %>%
  count()
## There are 2257 NAs
```

### Check values in Household number by age variables.
Check for anything >8.
```{r Value Household number by age variables}
ppv %>%
  select(ID, Household, HH_number, HH_u5:HH_2_45_64) %>%
  gather(HH_num_age, num, -ID:-HH_number) %>%
  filter(num>8) %>%
  spread(HH_num_age, num)
## There are 6 Household number by age entries that are >8. Comparing these with other Household variables suggests these can be explained and are genuine.
```

## Cross-check HH_number with Household numbers by age (HH-number-by-age) variables.

Compare HH_number with: 
1. Diagreements summing across HH-number-by-age variables
2. NAs


```{r HH_number - HH-number-by-age disagreements}
## Compare HH_number with sum of Household members by age.

## Disagreements including those caused by -99 in HH-number-by-age.
#ppv.tmp <- 
ppv %>%
  select(ID, Household, HH_number, HH_u5:HH_2_45_64) %>%
  gather(HH_num_age, num, -ID:-HH_number) %>%
  group_by(ID) %>%
  mutate(HH_sum = sum(num, na.rm=T)) %>%
  ungroup() %>%
  mutate(HH_num_age_dis = ifelse(HH_sum<0, 2, # -99 values
                          ifelse(HH_number!=HH_sum, 1, NA))) %>% # Real disagreements
  spread(HH_num_age, num) %>%
  filter(HH_num_age_dis%in%1:2) %>%
  View()
## 36 records 

## Disagreements caused by -99 values
ppv %>%
  select(ID, Household, HH_number, HH_u5:HH_2_45_64) %>%
  gather(HH_num_age, num, -ID:-HH_number) %>%
  group_by(ID) %>%
  mutate(HH_sum = sum(num, na.rm=T)) %>%
  ungroup() %>%
  mutate(HH_num_age_dis = ifelse(HH_sum<0, 2, # -99 values
                          ifelse(HH_number!=HH_sum, 1, NA))) %>% # Real disagreements
  spread(HH_num_age, num) %>%
  filter(HH_num_age_dis==2) %>%
  View()
## There are 16 records with -99 entries within all the ten Household numbers by age variables.


ppv %>%
  select(ID, Household, HH_number, HH_u5:HH_2_45_64) %>%
  gather(HH_num_age, num, -ID:-HH_number) %>%
  group_by(ID) %>%
  mutate(HH_sum = sum(num, na.rm=T)) %>%
  ungroup() %>%
  mutate(HH_num_age_dis = ifelse(HH_sum<0, 2, # -99 values
                          ifelse(HH_number!=HH_sum, 1, NA))) %>% # Real disagreements
  spread(HH_num_age, num) %>%
  filter(HH_num_age_dis==1) %>%
  View()
## There are 20 real disagreements between HH_numbers and HH-number-by-age.
```


Add flagging variable for HH_number - HH-number-by-age disagreements
```{r Add HH_number - HH-number-by-age disagreement flag}
## Produce df with ID and HH_num_age_dis (flagging HH_number - HH-number-by-age disagreements )
HH_age_check <- ppv %>%
  select(ID, Household, HH_number, HH_u5:HH_2_45_64) %>%
  gather(HH_num_age, num, -ID:-HH_number) %>%
  group_by(ID) %>%
  mutate(HH_sum = sum(num, na.rm=T)) %>%
  ungroup() %>%
  mutate(HH_num_age_dis = ifelse(HH_sum<0, 2, # -99 values
                          ifelse(HH_number!=HH_sum, 1, NA))) %>% # Real disagreements
  spread(HH_num_age, num) %>%
  filter(HH_num_age_dis%in%1:2) %>%
  select(ID, HH_num_age_dis)

## Temporary ppv df
ppv.tmp <- ppv

## Join ppv.tmp and HH_age_check using ID
ppv.tmp.j <- left_join(ppv.tmp, HH_age_check, "ID")

ppv.tmp.j %>%
  select(ID, HH_number, HH_u5:HH_2_65plus, HH_num_age_dis) %>%
  filter(HH_num_age_dis%in%1:2) %>%
  View()
## Verified

## Replace ppv
ppv <- ppv.tmp.j

## Clean up
rm(ppv.tmp, ppv.tmp.j, HH_age_check)
```

## HH_number merge single household number.

Instruction: Merge single person household numbers into here as these are missing.

Find single person Households that do not have HH_number=1.
```{r Find single person households}
ppv %>%
  select(ID, Household, HH_number, HH_text_code) %>%
  filter(Household==1, is.na(HH_number) | HH_number!=1) %>%
  View()
```
There are 870 'Single adult living alone' Household entries which don't have a comparable HH_number=1. These are all NAs.

```{r Update HH_number where Household is 1}
## IDs to update - for cross-checking
HH_singles <- ppv %>%
  filter(Household==1, is.na(HH_number) | HH_number!=1) %>%
  select(ID)

ppv.tmp <- ppv

## Update HH_number
ppv.tmp2 <- ppv.tmp %>%
  mutate(HH_number = ifelse(is.na(HH_number) & Household==1,1,HH_number))

## Cross-check update
ppv.tmp2 %>%
  anti_join(HH_singles) %>%
  View()
## Zero records returned

## Save update
ppv <- ppv.tmp2

## Clean up
rm(ppv.tmp, HH_singles, ppv.tmp2)
```


## Other HH_number NAs 
Check for other NAs in HH_number.

```{r Household value codes}
codes <- c(1:6,8)
text <- c("Single adult living alone",
"Single adult living with other adults", "A couple without children", "A couple with children", "Other", "Single adult living with children", "Refused to answer")
Household_v <- data.frame(Household=codes, Value=text)
rm(codes,text)
```

```{r Find other NAs in HH_number}
ppv %>%
  filter(!is.na(Household) & is.na(HH_number)) %>%
  group_by(Household) %>%
  summarise(count = n()) %>%
  right_join(Household_v, by = "Household") %>%
  select(Value, count, Household) %>%
  knitr::kable()
```
Need to update HH_number from the following:

|Value                                 | count| Household|
|:-------------------------------------|-----:|---------:|
|Single adult living with other adults |     2|         2|
|A couple without children             |  1376|         3|
|A couple with children                |     6|         4|
|Other                                 |     2|         5|
|Single adult living with children     |     1|         6|

### Update 'Single adult living with other adults'
There are 2 instances of Household=2 but HH_number is NA.
Is there any additional information to be able to determine the size of the household?

```{r Investigate HH_number where Household is 2}
ppv %>%
  filter(Household==2, is.na(HH_number)) %>%
  select(ID, HH_number, Household, HH_text, HH_u5:HH_2_65plus) %>%
  knitr::kable()
```
There is no additional information that can be used here to infer HH_number.
Create a variable to flag this issue.

```{r Add flagging variable HH_number_query}
ppv <- ppv %>%
  mutate(HH_number_query = ifelse(Household==2 & is.na(HH_number), 1, NA))
```


### 'A couple without children' NAs in HH_number:
There are 1376 instances of Household=3 (A couple without children) but HH_number is NA.
HH_number can be updated to 2 for these records.
```{r Update HH_number where Household is 3}
## IDs to update - for cross-checking
HH_coupleNoChild <- ppv %>%
  filter(Household==3, is.na(HH_number)) %>%
  select(ID)

ppv.tmp <- ppv

## Update HH_number
ppv.tmp2 <- ppv.tmp %>%
  mutate(HH_number = ifelse(is.na(HH_number) & Household==3,2,HH_number))

## Cross-check update
ppv.tmp2 %>%
  semi_join(HH_coupleNoChild) %>%
  View()

ppv.tmp2 %>% filter(HH_number==2 & Household==3) %>% select(ID, HH_number, Household)
## 1376 records returned

## Save update
ppv <- ppv.tmp2

## Clean up
rm(HH_coupleNoChild, ppv.tmp, ppv.tmp2)
```

### Update 'A couple with children'
There are 6 instances of Household=4 (A couple with children) but HH_number is NA.
Is there any additional information to be able to determine the size of the household?
```{r Investigate HH_number where Household is 4}
## 
ppv %>%
  filter(Household==4, is.na(HH_number)) %>%
  select(ID, HH_number, Household, HH_text, HH_u5:HH_2_65plus) %>%
  knitr::kable()
```
There appears to be no additional information that can be used to infer HH_number here.
Update HH_number_query variable flagging this issue.

```{r Update flagging variable HH_number_query}
ppv.tmp <- ppv

ppv.tmp <- ppv.tmp %>%
  mutate(HH_number_query = ifelse(Household==4 & is.na(HH_number), 1, HH_number_query))

ppv.tmp %>% filter(HH_number_query==1) %>% select(ID, Household, HH_number, HH_number_query)

ppv <- ppv.tmp

rm(ppv.tmp)
```


### Update 'Other'
There are 2 instances of Household=5 (Other) but *HH_number* is NA.
Is there any additional information to be able to determine the size of the household?
```{r Investigate HH_number where Household is 5}
ppv %>%
  filter(Household==5, is.na(HH_number)) %>%
  select(ID, HH_number, Household, HH_text, HH_u5:HH_2_65plus) %>%
  knitr::kable()
```
The HH_text for record ID 1009 indicates that this household consists of a 'couple with a child. So we can update HH_number for this record to 3.

Record ID 4231 contains no additional information that can be used to infer *HH_number*. *HH_number_query* variable will be updated to flag this issue.

```{r Update HH_number as 3 for ID 1009}
ppv.tmp <- ppv

ppv.tmp <- ppv.tmp %>%
  mutate(HH_number = ifelse(ID==1009, 3, HH_number))

ppv.tmp %>% filter(ID==1009) %>% select(ID, Household, HH_number)

ppv <- ppv.tmp

rm(ppv.tmp)
```


```{r Update flagging variable HH_number_query for ID 4231}
ppv.tmp <- ppv

ppv.tmp <- ppv.tmp %>%
  mutate(HH_number_query = ifelse(ID==4231, 1, HH_number_query))

ppv.tmp %>% filter(HH_number_query==1) %>% select(ID, Household, HH_number, HH_number_query)

ppv <- ppv.tmp

rm(ppv.tmp)
```

### 'Single adult living with children'
There is 1 instance of *Household*=6 (Single adult living with children) but *HH_number() is NA.
Is there any additional information to be able to determine the size of the household?

```{r Investigate HH_number where Household is 6}
ppv %>%
  filter(Household==6, is.na(HH_number)) %>%
  select(ID, HH_number, Household, HH_text, HH_u5:HH_2_65plus) %>%
  knitr::kable()
```
There appears to be no additional information that can be used to infer *HH_number* here.
Update *HH_number_query* variable flagging this issue.

```{r Update flagging variable HH_number_query}
ppv.tmp <- ppv

ppv.tmp <- ppv.tmp %>%
  mutate(HH_number_query = ifelse(Household==6 & is.na(HH_number), 1, HH_number_query))

ppv.tmp %>% filter(HH_number_query==1) %>% select(ID, Household, HH_number, HH_number_query)

ppv <- ppv.tmp

rm(ppv.tmp)
```









