---
title: "HH age option checks"
author: "Mark Ruddy"
date: "19 July 2016"
output: html_document
---

```{r Libraries}
library(dplyr)
library(ggplot2)
library(tidyr)
```

## Tasks

Sense check all these in relation to hh structure; flag anything odd.

## Sense check

### Initial exploration
```{r Preliminary data explortation}
ppv %>%
  select(ID, HH_u5:HH_2_65plus) %>%
  gather(AgeGrp, Number, -ID, na.rm=T) %>%
  filter(Number>0) %>%
  group_by(AgeGrp) %>%
  summarise(max=max(Number),
            min=min(Number),
            count=n()) %>%
  knitr::kable()
```

Minimum and maximum household number values by age:

|AgeGrp      | max| min| count|
|:-----------|---:|---:|-----:|
|HH_16_24    |   5| 0.5|   847|
|HH_2_16_24  |   2| 1.0|    37|
|HH_2_65plus |   2| 1.0|   266|
|HH_25_44    |  60| 1.0|  1472|
|HH_45_64    |  10| 1.0|  1435|
|HH_5_15     |   5| 1.0|   810|
|HH_65plus   |   6| 1.0|   273|
|HH_u5       |   3| 1.0|   503|


### Investigate not whole numbers.
```{r Outliers that are not whole numbers}
## eg in HH_16_24
## Function to test for whole numbers
library(gmp) # For is.whole() function
ppv %>%
  select(ID, HH_number, HH_u5:HH_2_65plus) %>%
  gather(AgeGrp, Number, -ID, na.rm=T) %>%
  mutate(int = is.whole(Number)) %>%
  filter(int==F) %>%
  knitr::kable()

```
Record ID 4811 has a value of *HH_16_24*=0.5 and HH_number=2.5

```{r Investigate record ID 4811 further}
glimpse(ppv %>% filter(ID==4811))
```
There is some evidence to suggest the value for *HH_16_24* here should be 1 and the *HH_number* value should be 3. This is because *Household*=4 (A couple with children). I will therefore update *HH_16_24* to 1 and *HH_number* to 3 for record ID=4811, assuming only one child is present in the household.

```{r Update HH_16_24 to 1 and HH_number to 3 for record ID 4811}
ppv.tmp <- ppv

## Update ID=4811
ppv.tmp <-ppv.tmp %>%
  mutate(HH_16_24 = ifelse(ID==4811, 1, HH_16_24)) %>%
  mutate(HH_number = ifelse(ID==4811, 3, HH_number))

## Check update
ppv.tmp %>% filter(ID%in%c(4811-10):c(4811+10)) %>% select(ID, HH_number, HH_u5:HH_2_65plus)

ppv <- ppv.tmp

## Clean up
rm(ppv.tmp)
```


### Investigate outliers
Exploring *HH-age-number* variable outliers as well as HH_number in more detail.
```{r Investigate records with outliers greater-than 9}
ppv %>%
  select(ID, HH_u5:HH_2_65plus) %>%
  gather(AgeGrp, Number, -ID, na.rm=T) %>%
  filter(Number>9) %>%
  spread(AgeGrp, Number) %>%
  knitr::kable()
```

|   ID| HH_25_44| HH_45_64|
|----:|--------:|--------:|
| 1213|       12|       NA|
| 1635|       12|       NA|
| 3213|       12|       NA|
| 3599|       60|       10|
| 5090|       11|       NA|

Record ID 3599 has 60 25-44 year-old adults and 10 45-64 year-old adults.

```{r Explore record ID 3599}
ppv %>%
  filter(ID==3599) %>%
  select(ID, HH_number, HH_text, HH_u5:HH_2_65plus)
```
This record comes from a response from an individual who lives in hostel (from HH_text) of 70 people. This record will be left unchanged but is highlighted as a potential outlier in later analysis.


```{r Explore records where HH_25_44 greater-than 10}
ppv %>%
  filter(HH_25_44>9 & HH_25_44<59) %>%
  select(ID, HH_number, HH_text, HH_u5:HH_2_65plus)
```
These records of Household sizes between 11 and 12 could all be considered valid - HH_number matches or HH_text supports.


```{r Plot household number values by age, excluding extremes}
## Excluding ID=3599 
ppv.HH_age <- ppv %>%
  select(ID, HH_u5:HH_2_65plus) %>%
  gather(AgeGrp, Number, -ID, na.rm=T) %>%
  filter(Number>0 & ID!=3599) %>%
  group_by(AgeGrp)
  
p.HH_age <- ggplot(ppv.HH_age, aes(AgeGrp, Number))

p.HH_age + 
  geom_jitter(height=0.2, width=0.6, alpha=0.3) + 
  scale_y_continuous(breaks=seq(1, 13, 1)) +
  labs(title="People per household age-group", x="Age Group", y="Number in household")

## Clean up
rm(Household_v, p.HH_age, ppv.HH_age)
```

The vast majority of values fall between 1 nd 5.  Investigate outliers in plot.
*HH_65plus*=6
```{r Investigate HH_65plus outlier in plot}
ppv %>% 
  filter(HH_65plus>5) %>%
  glimpse()
```
Nothing in information t

```{r Investigate HH_25_44 outliers in plot}
ppv %>% 
  filter(HH_25_44>5) %>%
  select(ID, HH_number, HH_text, HH_u5:HH_2_65plus, Borough, Source) %>%
  arrange(HH_number) %>%
  View()
```
Responses 793 and 1213 come from a shared house (*HH_text*). It is possible that ID 793 and 2441 come from the same household because they have the same age-number split in the same Borough. Although there is no other evidence supporting these higher *HH-age* and *HH_number* values within each record, there are no internal inconsistencies or counter evidence to suggest that these are erroneous.






















