---
title: "HH-Vehicle"
author: "Mark Ruddy"
date: "21 July 2016"
output: html_document
---


## Tasks

Variables: HH_carvan:HH_kidcyc ("HH_carvan", "HH_mcycle", "HH_othermv", "HH_adcyc", "HH_kidcyc")

1. Change refused to missing
2. Sense check; flag up anything odd
3. Create new variable for no car hh/one car hh/two or more car hh


## Change refused to missing

From 6.EntryChecks.Rmd and codebook.pdf there are no 'refused to answer' coded values.

## Sense checks

### Data type
HH_carvan:HH_kidcyc are character variables and need converting to numeric for best evaluaiton results.
```{r Convert to numeric}

ppv.tmp <- ppv

ppv.tmp2 <- ppv.tmp %>% 
  mutate(HH_carvan = as.numeric(HH_carvan)) %>%
  mutate(HH_mcycle = as.numeric(HH_mcycle)) %>%
  mutate(HH_othermv = as.numeric(HH_othermv)) %>%
  mutate(HH_adcyc = as.numeric(HH_adcyc)) %>%
  mutate(HH_kidcyc = as.numeric(HH_kidcyc))

## Check
ppv.tmp %>% 
  select(ID, HH_kidcyc) %>% # HH_carvan, HH_mcycle, HH_othermv, HH_adcyc, HH_kidcyc
  gather(v, n, -ID, na.rm=T) %>%
  group_by(n) %>%
  summarise(count=n())

ppv.tmp2 %>% 
  select(ID, HH_kidcyc) %>% # HH_carvan, HH_mcycle, HH_othermv, HH_adcyc, HH_kidcyc
  gather(v, n, -ID, na.rm=T) %>%
  group_by(n) %>%
  summarise(count=n())
## All ok

## Set changes
ppv <- ppv.tmp2

## Clean up
rm(ppv.tmp2, ppv.tmp)

```


```{r Preliminary data exploration}
ppv %>%
  select(ID, HH_carvan:HH_kidcyc) %>%
  gather(VehGrp, Number, -ID, na.rm=T) %>%
  filter(Number>0) %>%
  group_by(VehGrp) %>%
  summarise(max=max(Number),
            median=median(Number),
            min=min(Number),
            total=n()) %>%
  knitr::kable()
```
|VehGrp   | max| median| min| total|
|:----------|---:|------:|---:|-----:|
|HH_adcyc   |  16|      2|   1|  2292|
|HH_carvan  |  10|      1|   1|  3476|
|HH_kidcyc  |   8|      2|   1|   650|
|HH_mcycle  |  12|      1|   1|   238|
|HH_othermv |  10|      1|   1|    32|


```{r Plot HH transport to explore variation}

ppv.HH_veh <- ppv %>%
  select(ID, HH_carvan:HH_kidcyc) %>%
  gather(VehGrp, Number, -ID, na.rm=T) %>%
  filter(Number>0) %>%
  group_by(VehGrp)
  
p.HH_veh <- ggplot(ppv.HH_veh, aes(VehGrp, Number))

p.HH_veh + 
  geom_jitter(height=0.2, width=0.5, alpha=0.3) + 
  scale_y_continuous(breaks=seq(1, 17, 1)) +
  labs(title="Vehicle types in households", x="Vehicle type", y="Number in household")

## Clean up
rm(p.HH_veh, ppv.HH_veh)

```

Given the spread of values within each group in these data, a flagging variable *HH_vehicle_query* will be created for the following outliers.

HH_carvan > 4
HH_mcycle > 3
HH_othermv > 2
HH_adcyc > 6
HH_kidcyc > 4

This will allow open decision making later in the data analysis process.

```{r Create flagging variable HH_vehicle_query}
ppv.tmp <- ppv

ppv.tmp2 <- ppv.tmp %>%
  mutate(HH_vehicle_query = ifelse(HH_carvan>4, 1, NA)) %>%
  mutate(HH_vehicle_query = ifelse(is.na(HH_vehicle_query), ifelse(HH_mcycle>3, 1, HH_vehicle_query), HH_vehicle_query)) %>%
  mutate(HH_vehicle_query = ifelse(is.na(HH_vehicle_query), ifelse(HH_othermv>2, 1, HH_vehicle_query), HH_vehicle_query)) %>%
  mutate(HH_vehicle_query = ifelse(is.na(HH_vehicle_query), ifelse(HH_adcyc>6, 1, HH_vehicle_query), HH_vehicle_query)) %>%
  mutate(HH_vehicle_query = ifelse(is.na(HH_vehicle_query), ifelse(HH_kidcyc>4, 1, HH_vehicle_query), HH_vehicle_query))

## Check
ppv.tmp2 %>%
  select(ID, HH_vehicle_query, HH_carvan:HH_kidcyc) %>%
  View()
## OK

ppv %>% filter(HH_vehicle_query==1) %>% tally()

## Set changes
ppv <- ppv.tmp2

## Clean up
rm(ppv.tmp, ppv.tmp2)
```
There are 43 records that require vetting in terms of household vehicle numbers.


## Create new variable for no car hh/one car hh/two or more car hh
This is based upon variable HH_carvan.
Where 
```{r Create new variable HH_carvan_sum}

ppv.tmp <- ppv

ppv.tmp <- ppv.tmp %>%
  mutate(HH_carvan_sum = ifelse(is.na(HH_carvan), 0, 
                                ifelse(HH_carvan==0, 0,
                                       ifelse(HH_carvan==1, 1,
                                              ifelse(HH_carvan>1, 2, NA))))) 

## Check update
ppv.tmp %>%
  select(ID, HH_carvan, HH_carvan_sum) %>%
  sample_frac(0.1, replace = F) %>% # 10% check
  View()
## All ok

## Set changes
ppv <- ppv.tmp

## Clean up
rm(ppv.tmp)
```










