---
title: "Diary"
author: "Mark Ruddy"
date: "21 July 2016"
output: html_document
---


```{r Load packages}
library(dplyr)
library(tidyr)
library(stringr)
library(knitr)
```

## Tasks

1. Sense check and change refused to missing
2. Create new variables for any use of mcycle, PT, or taxi during that week
3. Create variables for how many days each mode was used (e.g. Dayswalking etc.)
4. Create new variables for 
  a) number of days' walking, 
  b) number of days cycling, 
  c) number of days active travel in past week

## 1. Sense check

There are no refused entries (see 6.EntryChecks.Rmd).

```{r View data}
ppv %>%
  select(ID, starts_with("Diary")) %>%
  sample_frac(0.1,replace = F) %>%
  View()
```


Possible issues:

### Are there any entries in Diary?none alongside other entries for that day?
```{r Diary?none with Diary?some check}
ppv.tmp <- ppv

## Tidy data and introduce 2 more IDs: one for each entry, one for each day entered per response
ppv.d.long <- ppv.tmp %>%
  select(ID, starts_with("Diary")) %>% 
  gather(Diary, Entry, -ID) %>%
  mutate(EntryID = 1:dim(.)[1], DayID = sprintf("%s%s", ID, substr(Diary, 1, 6))) %>%
  arrange(ID, Diary)

## Filter in 3 steps for where 'none' responses also have another response for the same day.
ppv.d.long %>%
  filter(grepl("none", Diary) & Entry==1) %>%
  inner_join(ppv.d.long, "DayID") %>%
  filter(!Entry.y==1 & !Diary.y==grepl("none", Diary.y)) %>%
  arrange(ID.y, Diary.y) %>%
  filter(Entry.y==1)
  View()

## There aren't any!

## Clean up
rm(ppv.tmp, ppv.d.long)
```
No conflicts between Diary?none and other diary entries.


### Are there daily totals of different modes that seem implausible?

Within records, check how many modes were entered for each day.

```{r Modes per day check}
ppv.tmp <- ppv

## Check sum entries per day within IDs
diary <- c("Diary1", "Diary2", "Diary3", "Diary4", "Diary5", "Diary6", "Diary7")
ppv.tmp %>% select(starts_with(diary[1])) %>% # replace index in diary[1] to choose day
  rowSums(na.rm=T) %>%
  data.frame(.) %>% 
  bind_cols(ppv.tmp, .) %>%
  setNames(., c(names(ppv.tmp), "DaySum")) %>%
  filter(DaySum>=4) %>% 
  select(ID, DaySum, starts_with(diary[1])) %>% # replace index in diary[1] to choose day
  gather(DiaryDay, DaySum, -ID) %>%
  arrange(desc(DaySum)) %>%
  View()

## Clean up
rm(diary, ppv.tmp)
```
Some days have up to 6 modes entered. These seem to be valid but a flag will be created to allow assessment later in the project.

Create flagging variable.
```{r Create flag for 5 or more modes per day}
## New group_by approach
ppv.tmp <- ppv

## find responses where there are 5 or more modes entered for a day
modeGT4 <- ppv.tmp %>%
  select(ID, starts_with("Diary")) %>% 
  gather(Diary, Entry, -ID) %>%
  mutate(Diary.day = substr(Diary, 1, 6)) %>%
  group_by(ID, Diary.day) %>%
  summarise(Sum=sum(Entry, na.rm=T)) %>%
  ungroup() %>%
  filter(Sum>=5) %>%
  distinct(ID) %>%
  select(ID, Sum)

## Create flagging variable 
ppv.tmp <- ppv.tmp %>%
  mutate(Diary_Sum_Query = ifelse(ID%in%unlist(modeGT4[,"ID"]), 1, NA)) 

## Check new variable
ppv.tmp %>%
  select(ID, starts_with("Diary"), Diary_Sum_Query) %>% 
  filter(Diary_Sum_Query==1) %>%
  View()

## Set changes
ppv <- ppv.tmp

## Clean up
rm(modeGT4, ppv.tmp)
```
There are 9 responses where there are 5 or more modes listed in one day.



## 2. Create new variable(s?) for any use of mcycle, PT, or taxi during that week
```{r Create mcycle, PT, or taxi variable}
## Temporary dataset
ppv.tmp <- ppv

# Total cyc
totModes <- ppv.tmp %>%
  select(ID, ends_with("PT"), ends_with("taxi"), ends_with("mcyc")) %>% 
  select(ID, starts_with("Diary")) %>%
  mutate(Total_mode = rowSums(.[,-1], na.rm=T)) %>%
  mutate(Total_mode = ifelse(Total_mode==0, NA, Total_mode)) %>%
  arrange(Total_mode) %>%
  select(ID, Total_mode) #%>% View()


# Join Total_cyc variable to dataset
ppv.tmp2 <- ppv.tmp %>%
  left_join(totModes, "ID") %>%
  mutate(Total_mode = ifelse(!is.na(Total_mode), 1, NA)) %>%
  rename(Total_mcycTaxiPT = Total_mode)

## Set changes
ppv <- ppv.tmp2
  
## Clean up
rm(ppv.tmp, ppv.tmp2, totModes)
```


## 3. Variables for how many days each mode was used

### Walking
How many total days walking from Diary?walk
```{r Total diary days walking}
## Temporary dataset
ppv.tmp <- ppv

# Total walk
totWalk <- ppv.tmp %>%
  select(ID, starts_with("Diary")) %>%
  select(ID, ends_with("walk")) %>%
  mutate(Total_walk = rowSums(.[,-1], na.rm=T)) %>%
  mutate(Total_walk = ifelse(Total_walk==0 & is.na(rowSums(.[,-1])), NA, Total_walk)) %>%
  select(ID, Total_walk)

# Join Total walk variable to dataset
ppv.tmp2 <- ppv.tmp %>%
  left_join(totWalk, "ID") 

#glimpse(ppv.tmp2)

## Check
# ppv.tmp2 %>%
#   select(ID, starts_with("Diary"), Total_walk) %>%
#   select(ID, ends_with("walk")) %>%
#   sample_frac(0.1) %>%
#   arrange(Total_walk) %>%
#   View()
## OK

## Set changes
ppv <- ppv.tmp
  
## Clean up
rm(ppv.tmp, ppv.tmp2, totWalk)
```


### Cycles
How many total days cycling from Diary?cyc
```{r Total diary days cycling}
## Temporary dataset
ppv.tmp <- ppv

# Total cyc
totCyc <- ppv.tmp %>%
  select(ID, starts_with("Diary")) %>%
  select(ID, ends_with("cyc")) %>%
  select(-ends_with("mcyc")) %>% # because 'cyc' picks up 'mcyc' too so need to remove latter
  mutate(Total_cyc = rowSums(.[,-1], na.rm=T)) %>%
  mutate(Total_cyc = ifelse(Total_cyc==0 & is.na(rowSums(.[,-1])), NA, Total_cyc)) %>%
  select(ID, Total_cyc)
  #sample_frac(0.1) %>% arrange(Total_cyc) # 10% check

# Join Total_cyc variable to dataset
ppv.tmp2 <- ppv.tmp %>%
  left_join(totCyc, "ID") 

# glimpse(ppv.tmp2)

## Check
# ppv.tmp2 %>%
#   select(ID, starts_with("Diary"), Total_cyc) %>%
#   select(ID, ends_with("cyc")) %>%
#   select(-ends_with("mcyc")) %>%
#   sample_frac(0.1) %>%
#   arrange(Total_cyc) %>%
#   View()
## OK

## Set changes
ppv <- ppv.tmp2
  
## Clean up
rm(ppv.tmp, ppv.tmp2, totCyc)
```


### Motorcycles
How many total days motorcycling from Diary?mcyc
```{r Total diary days motorcycling}
## Temporary dataset
ppv.tmp <- ppv

# Total mcyc
totMcyc <- ppv.tmp %>%
  select(ID, starts_with("Diary")) %>%
  select(ID, ends_with("mcyc")) %>%
  mutate(Total_mcyc = rowSums(.[,-1], na.rm=T)) %>%
  mutate(Total_mcyc = ifelse(Total_mcyc==0 & is.na(rowSums(.[,-1])), NA, Total_mcyc)) %>%
  select(ID, Total_mcyc)

# Join Total_mcyc variable to dataset
ppv.tmp2 <- ppv.tmp %>%
  left_join(totMcyc, "ID") 

# glimpse(ppv.tmp2)

## Check
# ppv.tmp2 %>%
#   select(ID, starts_with("Diary"), Total_mcyc) %>%
#   select(ID, ends_with("mcyc")) %>%
#   sample_frac(0.1) %>%
#   arrange(Total_mcyc) %>%
#   View()
## OK

## Set changes
ppv <- ppv.tmp2
  
## Clean up
rm(ppv.tmp, ppv.tmp2, totMcyc)
```


### Public transport
How many total days public transport from Diary?mcyc
```{r Total diary days public transport}
## Temporary dataset
ppv.tmp <- ppv

# Total PT
totPT <- ppv.tmp %>%
  select(ID, starts_with("Diary")) %>%
  select(ID, ends_with("PT")) %>%
  mutate(Total_PT = rowSums(.[,-1], na.rm=T)) %>%
  mutate(Total_PT = ifelse(Total_PT==0 & is.na(rowSums(.[,-1])), NA, Total_PT)) %>% 
  select(ID, Total_PT)
  

# Join Total_PT variable to dataset
ppv.tmp2 <- ppv.tmp %>%
  left_join(totPT, "ID") 

# glimpse(ppv.tmp2)

## Check
# ppv.tmp2 %>%
#   select(ID, starts_with("Diary"), Total_PT) %>%
#   select(ID, ends_with("PT")) %>%
#   sample_frac(0.1) %>%
#   arrange(Total_PT) %>%
#   View()
## OK

## Set changes
ppv <- ppv.tmp2
  
## Clean up
rm(ppv.tmp, ppv.tmp2, totPT)
```

### Taxi
How many total days using taxis from Diary?mcyc
```{r Total diary days taxi}
## Temporary dataset
ppv.tmp <- ppv

# Total taxi
totTaxi <- ppv.tmp %>%
  select(ID, starts_with("Diary")) %>%
  select(ID, ends_with("taxi")) %>%
  mutate(Total_taxi = rowSums(.[,-1], na.rm=T)) %>%
  mutate(Total_taxi = ifelse(Total_taxi==0 & is.na(rowSums(.[,-1])), NA, Total_taxi)) %>% 
  select(ID, Total_taxi)
  

# Join Total_taxi variable to dataset
ppv.tmp2 <- ppv.tmp %>%
  left_join(totTaxi, "ID") 

# glimpse(ppv.tmp2)

## Check
# ppv.tmp2 %>%
#   select(ID, starts_with("Diary"), Total_taxi) %>%
#   select(ID, ends_with("taxi")) %>%
#   sample_frac(0.1) %>%
#   arrange(Total_taxi) %>%
#   View()
## OK

## Set changes
ppv <- ppv.tmp2
  
## Clean up
rm(ppv.tmp, ppv.tmp2, totTaxi)
```

### Car
How many total days driving from Diary?mcyc
```{r Total diary days driving}
## Temporary dataset
ppv.tmp <- ppv

# Total car
totCar <- ppv.tmp %>%
  select(ID, starts_with("Diary")) %>%
  select(ID, ends_with("car")) %>%
  mutate(Total_car = rowSums(.[,-1], na.rm=T)) %>%
  mutate(Total_car = ifelse(Total_car==0 & is.na(rowSums(.[,-1])), NA, Total_car)) %>% 
  select(ID, Total_car)
  

# Join Total_car variable to dataset
ppv.tmp2 <- ppv.tmp %>%
  left_join(totCar, "ID") 

# glimpse(ppv.tmp2)

## Check
# ppv.tmp2 %>%
#   select(ID, starts_with("Diary"), Total_car) %>%
#   select(ID, ends_with("car")) %>%
#   sample_frac(0.1) %>%
#   arrange(Total_car) %>%
#   View()
## OK

## Set changes
ppv <- ppv.tmp2
  
## Clean up
rm(ppv.tmp, ppv.tmp2, totCar)
```

## 4. Create new variables for number of days' walking, cycling, active travel in past week

```{r Create walking-cycling variable}
## Temporary dataset
ppv.tmp <- ppv

## Gather cycling and walking diary entries. Generate combined DayModeID (ID+'diary day' - eg Diary1). DayModeID will be used to group each day's active travel.
act_long <- ppv.tmp %>%
  select(ID, ends_with("walk"), ends_with("cyc")) %>% 
  select(-ends_with("mcyc")) %>%
  select(ID, starts_with("Diary")) %>%
  gather(DayMode, Entry, -ID) #%>% View()

## Group active travel using DayModeID and sum.
act_days <- act_long %>%
  mutate(DayModeID = paste0(ID, substr(DayMode, 0, 6))) %>%
  group_by(DayModeID) %>%
  summarise(SumDayMode = sum(Entry, na.rm=T)) %>%
  ## Create 0,1 flag for an active day
  mutate(Active_flag = ifelse(SumDayMode>=1, 1, 0)) %>% ## There are no NAs in the results
  ## Recreate ID to group by for sum number days active
  mutate(ID = substr(DayModeID, 0, nchar(DayModeID)-6)) %>% mutate(ID = as.integer(ID)) %>%
  ## Group by ID and sum Active_day
  ungroup() %>% group_by(ID) %>%
  summarise(Active_days = sum(Active_flag))

## Test and check data, and introduce NAs
act_days_NA <- ppv.tmp %>%
  left_join(act_days, "ID") %>%
  select(ID, Active_days, ends_with("walk"), ends_with("cyc")) %>% 
  select(-ends_with("mcyc")) %>%
  select(ID, Active_days, starts_with("Diary")) %>%
  ## Make Active_days NA where all active mode diary entries NA
  mutate(Active_days = ifelse(Active_days==0 & is.na(rowSums(.[,-1])), NA, Active_days)) %>%
  select(ID, Active_days)
  #sample_frac(0.1) %>% 
  #arrange(Active_days) %>% View()

## Join to main ppv dataset by ID
ppv.tmp2 <- ppv.tmp %>%
  left_join(act_days_NA, "ID")

#glimpse(ppv.tmp2)

## Check
# ppv.tmp2 %>%
#   select(ID, Active_days, ends_with("walk"), ends_with("cyc")) %>% 
#   select(-ends_with("mcyc")) %>%
#   select(ID, Active_days, starts_with("Diary")) %>%
#   sample_frac(0.1) %>%
#   arrange(Active_days) %>%
#   View()
## OK

## Set changes
ppv <- ppv.tmp2
  
## Clean up
rm(ppv.tmp, ppv.tmp2, act_long, act_days, act_days_NA)
```








