---
title: "ModeDurationDay"
author: "Mark Ruddy"
date: "25 July 2016"
output: html_document
---

Load libraries and data
```{r Setup, include=FALSE}
library(dplyr)
library(tidyr)
library(stringr)
library(knitr)
library(ggplot2)
setwd("/Users/markruddy/Documents/Personal/Work/DataScience/PeopleAndPlaces/")
load("./Data/pp.data.RData")
```


## Tasks

Sense check and change refused to missing for Walk?min, Cyc?min and Car?min.
Look at daily entries of 8 hours (480 minutes) or more.

```{r Tidy data}

ppv.tmp <- ppv

## Tidy and group into modes
daymode_long <- ppv.tmp %>%
  select(ID, ends_with("min")) %>%
  gather(DayMode, Entry, -ID) %>%
  mutate(Mode = ifelse(grepl("Walk", DayMode), "Walking",
                       ifelse(grepl("Cyc", DayMode), "Cycling",
                              ifelse(grepl("Car", DayMode), "Driving", NA)))) %>%
  group_by(Mode)

## Look for records of activity of 8 hours or more
daymode_long %>% 
  select(ID, DayMode, Entry) %>%
  filter(Entry>=480) %>%
  ungroup() %>%
  arrange(desc(Entry)) %>%
  rename(Minutes=Entry) %>%
  kable(caption="List of responses with walking, cycling or driving of 8+ hours in a day")

## Count responses
daymode_long %>% 
  select(ID, DayMode, Entry) %>%
  filter(Entry>=480) %>%
  ungroup() %>%
  select(ID, -DayMode, -Entry) %>%
  group_by(ID) %>%
  summarise(count=n())
## 53 responses

## Check IDs
# ppv.tmp %>%
#   select(ID, Walk1min:Car7min) %>%
#   filter(ID==271) %>%
#   View()
```

* There are 99 entries from 53 responses of 8 or more hours per day for each activity.
* Seventeen entries are greater than 24 hours (1440 minutes) and obviously erroneous.
* Three entries are below 24 hours but above 14 hours and need further assessment.
* The remaining 79 entries may need further assessment.


## Plots of minutes entered by mode
Explore the distribution of entries for each mode with evidently erroneous values (above 840 minutes - 14hrs)) removed.
```{r Jitter plot minutes entered by mode}
## Remove error entries - above 840 minutes (14hrs) - and NAs
daymode_ <- daymode_long %>%
  na.omit() %>%
  filter(Entry<840) %>%
  mutate(Hours = Entry/60)

p.daymode <- ggplot(daymode_, aes(Mode, Hours))

p.daymode + 
  geom_jitter(height=0.2, width=0.5, alpha=0.3) + 
  #ylim(0, 90) +
  #ylim(0, 14) +
  scale_y_continuous(breaks=seq(0, 14, 1)) +
  labs(title="Minutes spent each day walking, cycling and driving", x="Mode", y="Hours")

## Clean up
rm(daymode_, p.daymode)

##? Possibility of a shiny dashboard solution here
```


## Add flagging variables for modes entered as 8 hours or more.

A flagging variable will be created for each mode indicating entries that are or exceed 8 hours (480 minutes).

```{r Add Mins greater or equal to 10 hours flagging variable}

ppv.tmp <- daymode_long %>%
  ungroup() %>%
  filter(Mode=="Walking") %>%
  mutate(Walk_min_query = ifelse(Entry>=480, 1, NA)) %>%
  spread(DayMode, Entry) %>%
  select(ID, Walk_min_query) %>%
  distinct(ID) %>%
  # filter(Walk_min_query==1)
  left_join(ppv.tmp, ., "ID")
  
ppv.tmp <- daymode_long %>%
  ungroup() %>%
  filter(Mode=="Cycling") %>%
  mutate(Cyc_min_query = ifelse(Entry>=480, 1, NA)) %>%
  spread(DayMode, Entry) %>%
  select(ID, Cyc_min_query) %>%
  distinct(ID) %>%
  left_join(ppv.tmp, ., "ID")

ppv.tmp <- daymode_long %>%
  ungroup() %>%
  filter(Mode=="Driving") %>%
  mutate(Car_min_query = ifelse(Entry>=480, 1, NA)) %>%
  spread(DayMode, Entry) %>%
  select(ID, Car_min_query) %>%
  distinct(ID) %>%
  # filter(Walk_min_query==1)
  left_join(ppv.tmp, ., "ID")

ppv.tmp %>%
  select(ID, Walk1min:Car7min, Walk_min_query, Cyc_min_query, Car_min_query) %>%
  filter(Walk_min_query==1 | Cyc_min_query==1 | Car_min_query==1)

## Set changes
ppv <- ppv.tmp

## Clean up
rm(daymode_long, ppv.tmp)
```

## Check for mis-matches between Diary and modal Mins entries.
Are there disagreements between days where diary entries (Diary?) and minutes spent (?min) entries

```{r Diary and modal Mins mis-matches}
## Test dataset
ppv.tmp <- ppv

## Select Diary? ?min variables of interest
diary.mins <- ppv.tmp %>%
  select(ID, starts_with("Diary"), ends_with("min")) %>%
  select(-contains("mcyc"), -contains("Query"), -ends_with("PT"), -ends_with("taxi"), -ends_with("unsure"), -ends_with("none"), -ends_with("ref")) #%>% glimpse()

## Tidy data and create new variable DayMode (just day and mode).
## This gives a value shared between Diary? and ?min entries for same day in a response that can be grouped-by later in mis-match test
diary.mins <- diary.mins %>%
  gather(Variable, Value, -ID) %>%
  mutate(DayMode = ifelse(grepl("Diary", Variable), substr(Variable, 6, nchar(Variable)), NA)) %>%
  mutate(DayMode = ifelse(grepl("^Walk", Variable), paste0(substr(Variable, 5, 5), "walk"), DayMode)) %>%
  mutate(DayMode = ifelse(grepl("^Cyc", Variable), paste0(substr(Variable, 4, 4), "cyc"), DayMode)) %>%
  mutate(DayMode = ifelse(grepl("^Car", Variable), paste0(substr(Variable, 4, 4), "car"), DayMode))


## Mis-match test
## Group-by ID and DatMode to check for consistency between two different entries for same day.
diary.mins <- diary.mins %>%
  mutate(Value = as.numeric(Value))

glimpse(diary.mins)

## Using dev version of dplyr July 27, 2016 from GitHub
diary.match <- diary.mins %>%
  group_by(ID, DayMode) %>%
  mutate(mismatch = ifelse(is.na(Value) & lag(is.na(Value)), 0, # Capture both NA
                       ifelse((is.na(Value) & lag(Value==0)) | (Value==0 & lag(is.na(Value))), 0, 
                          ifelse((Value>0 & lag(Value)==0) | (Value==0 & lag(Value)>0), 1, 
                             ifelse((Value>0 & lag(is.na(Value))) | (is.na(Value) & lag(Value>0)), 1, 0))))) #%>% View()

## Show which variables are in conflict
diary.mismatch <- diary.match %>% 
  ungroup() %>%
  filter(mismatch==1) %>%
  select(-Value, -mismatch) %>%
  arrange(ID, DayMode, Variable) %>%
  rename("ModeXmin" = Variable, "DiaryXmode" = DayMode) %>%
  mutate(DiaryXmode = paste0("Diary", DiaryXmode))  #%>%
  # kable(caption="DiaryXmode and ModeXmin entries in conflict")

## Check conflicts
## Matrix of mismatches
diary.mismatch.wide <- diary.mismatch %>%
  spread(DiaryXmode, ModeXmin) %>%
  setNames(paste0("Diary", names(.)))

## Check matrix to confirm presence of mismatches
ppv.tmp %>%
  select(ID, starts_with("Diary"), ends_with("min")) %>%
  select(-contains("mcyc"), -contains("Query"), -ends_with("PT"), -ends_with("taxi"), -ends_with("unsure"), -ends_with("none"), -ends_with("ref")) %>%
  filter(ID==267) %>%
  View()

rm(diary.mins, diary.match, ppv.tmp)
```
There are 181 responses where differences exist between DiaryXmode and ModeXmin entries for the same day. These can be found listed in table "diary.mismatch" and as a matrix in "diary.mismatch.wide".















