---
title: "Cycling behaviour variables"
author: "Mark Ruddy"
date: "27 July 2016"
output: html_document
---

Load libraries and data
```{r Setup, include=FALSE}
library(dplyr)
library(tidyr)
library(stringr)
library(knitr)
library(ggplot2)
# setwd("/Users/markruddy/Documents/Personal/Work/DataScience/PeopleAndPlaces/")
# load("./Data/pp.data.RData")
```

## Tasks

1. *Cancyc*: Create new variable where people who do cycle are added in (this was only asked to those who hadn't cycled in last week)
2. Create new variable adding in people who can't cycle (as not having cycled)
3. Combine *Other_ad_cyc* and *Other_ad_cyc2*
4. *HH_child_cyc*: Ensure missing (for ppl with no kids) is distinguished from any missing (did not answer)
5. *Commcyc*, *Esccyc*, *Foodcyc*, *Friendcyc*: Add in system missing and ensure distinguished from did not answer




## 1. *Cancyc* plus other that ride bicycles

Combine *Cancyc* with other indicators of having cycled in the last week (ie DairyXcyc and CycXmin), with Cycleweek and MinsCycling, and Cyc12mon and Cycmonth.

```{r Check extra people that ride a bicycle}
## Temp df
ppv.tmp <- ppv

## Select variables that indicate cycling use
cyc.dat <- ppv.tmp %>%
  select(ID, Cycleweek, MinsCycling, starts_with("Diary"), ends_with("min"), Cyc12mon, Cycmonth) %>%
  select(ID, MinsCycling, ends_with("cyc"), starts_with("Cyc"), -contains("mcyc")) %>% View()

## Create Cancyc2 variable showing evidence of cycle use from other cycling variables
cyc.can <- cyc.dat %>%
  mutate(Cancyc2 = ifelse(rowSums(.[2:17], na.rm=T)>0,1,NA)) %>% 
  mutate(Cancyc2 = ifelse(is.na(Cancyc2) & Cyc12mon==1, 1, ifelse(is.na(Cancyc2) & Cycmonth==1, 1, Cancyc2)))

## Check Cancyc2 records
cyc.can %>%
  filter(Cancyc2==1) %>% View()

## Cross-check Cancyc (!=1) with Cancyc2 (=1) for any missing cyclists
cyc.can %>%
  select(ID, Cancyc2) %>%
  left_join(ppv.tmp, ., "ID") %>%
  filter((Cancyc!=1 | is.na(Cancyc)) & Cancyc2==1) %>%
  # filter(Cancyc==1 | Cancyc2==1) %>%
  select(ID, Cycleweek, MinsCycling, starts_with("Diary"), ends_with("min"), Cyc12mon, Cycmonth, Cancyc, Cancyc2) %>%
  select(ID, Cycleweek, MinsCycling, Cancyc, Cancyc2, ends_with("cyc"), starts_with("Cyc"), -contains("mcyc")) %>%
  # select(ID, Cancyc, Cancyc2) %>%
  View() 
```
There are 923 additional entries for people that can ride a bicycle.

```{r Create new variable for all people who can ride a bicycle}
## New variable Cancyc_all that combines Cancyc and other evidence of cycling
ppv.tmp2 <- cyc.can %>%
  select(ID, Cancyc2) %>%
  left_join(ppv.tmp, ., "ID") %>%
  mutate(Cancyc_all = ifelse(is.na(Cancyc2) & Cancyc==1, 1, Cancyc2)) %>%
  select(-Cancyc2)

## Check update
ppv.tmp2 %>%
  # summarise(sum=sum(Cancyc_all, na.rm=T))
  select(ID, Cancyc, Cancyc_all, Cycleweek, MinsCycling, starts_with("Diary"), ends_with("min")) %>%
  select(ID, Cancyc_all, Cancyc, Cycleweek, MinsCycling, ends_with("cyc"), starts_with("Cyc"), -contains("mcyc")) %>%
  sample_n(200) %>%
  View()
## Update OK - now 3997 responses from people that can cycle

## Set changes
ppv <- ppv.tmp2

## Clean up
rm(cyc.can, cyc.dat, ppv.tmp, ppv.tmp2)
```
There are 3997 responses from people that can cycle in *Cancyc_all*


## 2. Create new variable adding in people who can't cycle (as not having cycled)

Need new variables that combines values of 2 and 3 (not able to ride a bicycle) from *Cancyc* with 'not ridden pedal cycle' responses from *Cyc12mon* and *Cycmonth*. 

Create new variables:
*Cycmonth_all_not*
*Cyc12mon_all_not*

Create *Cycmonth_all_not*
```{r Create Cycmonth_all_not}
## Temp dataset
ppv.tmp <- ppv

## Create new variable
ppv.tmp2 <- ppv.tmp %>%
  mutate(Cycmonth_all_not = ifelse(Cancyc %in% c(2,3), 1, 
                               ifelse(Cycmonth == 2, 1, NA)))

## Check
ppv.tmp2 %>% 
  select(ID, Cancyc, Cycmonth, Cycmonth_all_not) %>%
  # filter(Cycmonth_all_not==1) %>%
  # tally()
  sample_n(300) %>%
  View()
## OK

## Set changes
ppv <- ppv.tmp2

## Clean up
rm(ppv.tmp, ppv.tmp2)
```
Including those people who cannot cycle, there are 1389 records of not cycling in the last month.

Create *Cyc12mon_all_not*
```{r Create Cyc12mon_all_not}
## Temp dataset
ppv.tmp <- ppv

## Create new variable
ppv.tmp2 <- ppv.tmp %>%
  mutate(Cyc12mon_all_not = ifelse(Cancyc %in% c(2,3), 1, 
                               ifelse(Cyc12mon == 2, 1, NA)))
## Check
ppv.tmp2 %>% 
  select(ID, Cancyc, Cyc12mon, Cyc12mon_all_not) %>%
  # filter(Cyc12mon_all_not==1) %>%
  # tally()
  sample_n(200) %>%
  View()

# glimpse(ppv.tmp2)
## OK

## Set changes
ppv <- ppv.tmp2

## Clean up
rm(ppv.tmp, ppv.tmp2)
```
Including those people who cannot cycle, there are 2350 records of not cycling in the last 12 months.


## 3. Combine *Other_ad_cyc* and *Other_ad_cyc2*
Combine entries from these two variables.

```{r Combine Other_ad_cyc with Other_ad_cyc2}
## temp data
ppv.tmp <- ppv

## Create combined variable
ppv.tmp <- ppv.tmp %>%
  mutate(Other_ad_cyc_all = ifelse(Other_ad_cyc==1, 1, NA)) %>%
  mutate(Other_ad_cyc_all = ifelse(is.na(Other_ad_cyc_all), ifelse(Other_ad_cyc2==1, 1, Other_ad_cyc_all), Other_ad_cyc_all))

## Check
ppv.tmp %>%
  select(ID, Other_ad_cyc, Other_ad_cyc2, Other_ad_cyc_all) %>%
  sample_n(200) %>%
  View()
## OK

# ppv.tmp %>%
#   filter(Other_ad_cyc_all==1) %>%
#   tally()

## Set changes
ppv <- ppv.tmp

## Clean up
rm(ppv.tmp)
```
There are 2076 responses where respondant knows smeone who cycles regularly.


## 4. *HH_child_cyc*: Ensure missing (for ppl with no kids) is distinguished from any missing (did not answer)

### Sense check
Compare *HH_child_cyc* with indicators of respondant with children through *Household* values 4 and 6 and create new variable *HH_child_cyc_query* flagging conflicts
```{r HH_child_cyc compare with indicators of children}
## Test data
ppv.tmp <- ppv

## Select columns and filter for chcildren that cycle (HH_child_cyc==1 | HH_child_cyc==2) compared with possible indicator of no children (Household!=4 & Household!=6)
## Create new variable HH_child_cyc_query flagging conflicts
HHccq <- ppv.tmp %>% 
  select(ID, Household, HH_child_cyc, HH_text) %>%
  # filter(Household==4 | Household==6) %>%
  filter((HH_child_cyc==1 | HH_child_cyc==2) & (Household!=4 & Household!=6)) %>%
  mutate(HH_child_cyc_query = 1) %>%
  select(ID, HH_child_cyc_query) 

## Join with original table 
ppv.tmp2 <- ppv.tmp %>%
  full_join(HHccq, "ID") %>%
  glimpse()

## Check
ppv.tmp2 %>%
  select(ID, HH_child_cyc_query, Household, HH_child_cyc, HH_text) %>%
  View()
## OK

## Set changes
ppv <- ppv.tmp2

## Clean up
rm(ppv.tmp2, ppv.tmp, HHccq)
```
There are 30 responses where there is posible conflict between Household value and *HH_child_cyc* value. These are flagged in the *HH_child_cyc_query* variable.



### Ensure missing (for ppl with no kids) is distinguished from any missing (did not answer)

Set *HH_child_cyc* values for respondants without children to 5. Values that are NA mean there was no answer.

An categorical indictor of respondants without children is *Household* values 1 and 3.
```{r Set HH_child_cyc without children to -99}
## Test data
ppv.tmp <- ppv

## Check for any HH_child_cyc entries where Household is childless (Household==1 | Household==3).
ppv.tmp %>%
  select(ID, Household, HH_child_cyc, HH_text) %>%
  filter(!is.na(HH_child_cyc) & (Household==1 | Household==3)) %>%
  View()
## There are none of these

## Query for no HH_child_cyc entries in Households that are childless (Household==1 | Household==3). Update HH_child_cyc entries to 5 for this combination.
ppv.tmp2 <- ppv.tmp %>%
  mutate(HH_child_cyc = ifelse((Household==1 | Household==3), 5, HH_child_cyc))
## There are 2246 of these entries

## Check
glimpse(ppv.tmp2)

ppv.tmp2 %>%
  filter(HH_child_cyc==5) %>%
  tally()

ppv %>%
  select(ID, Household, HH_child_cyc) %>%
  sample_n(400) %>%
  arrange(desc(HH_child_cyc)) %>%
  View()
## OK

## Set changes
ppv <- ppv.tmp2

## Clean up
rm(ppv.tmp2,ppv.tmp)
```
Updated 2246 *HH_child_cyc* entries=5 to distingiuish childless households (*Household*=1 or 3) from missing *HH_child_cyc* entries.


## 5. *Commcyc*, *Esccyc*, *Foodcyc*, *Friendcyc*: Add in system missing and ensure distinguished from did not answer

Compare against other indicators of cycling.
```{r Scope variables}
ppv.tmp <- ppv

ppv.tmp %>%
  select(Commcyc) %>%
  group_by(Commcyc) %>%
  summarise(count=n())

## Check *Commcyc*, *Esccyc*, *Foodcyc* *Friendcyc* entries against other cycling indicators (*Cycevid*).
Reasoncyc <- ppv.tmp %>%
  select(ID, contains("Cycling"), starts_with("Cyc"), ends_with("cyc"), -contains("HH"), -contains("ad"), -contains("not"), -contains("Area"), -contains("Total"), -contains("query"), -Cancyc, -contains("mcyc"), Commcyc) %>%
  mutate(Cycevid = ifelse(rowSums(.[2:17], na.rm=T)>0,1,NA)) %>% 
  mutate(Cycevid = ifelse((is.na(Cycevid) & Cycmonth==1), 1, Cycevid)) 

## Where reason for cycling not alongside other cycling evidence
Reasoncyc %>%
  filter(Cycevid!=1 & (Commcyc>=1 | Esccyc>=1 | Foodcyc>=1 | Friendcyc>=1))
## There are no conflicts where *Commcyc*, *Esccyc*, *Foodcyc* *Friendcyc* are not also found with other indicators of cycling

## Clean up
rm(Reasoncyc, ppv.tmp)
```
There are no conflicts where *Commcyc*, *Esccyc*, *Foodcyc*, *Friendcyc* are not also found with other indicators of cycling.



















