---
title: "Cycling area"
author: "Mark Ruddy"
date: "29 July 2016"
output: html_document
---

Load libraries and data
```{r Setup, include=FALSE}
library(dplyr)
library(tidyr)
library(stringr)
library(knitr)
library(ggplot2)
# setwd("/Users/markruddy/Documents/Personal/Work/DataScience/PeopleAndPlaces/")
# load("./Data/pp.data.RData")
```

## Tasks

Create new variables combining agree and strongly agree; disagree and strongly disagree for variables:
*Area_cyc_unsafe*, *Area_cross*, *Area_talk*, *Area_8yo_walk*, *Area_stop*, *Area_crime*, *Area_cyc_path*, *Area_walk_pleas*, *Area_shade*, *Area_airpoll*, *Area_cyc_pleas*, *Area_8yo_cyc*, *Area_facilities*, *Area_walk_unsafe*, *Area_pavements*.

The values coding for these options are as follows:

Combine disagree and strongly disagree into 'Disagree'
'Tend to Disagree'=2
'Strongly Disagree'=1

Combine agree and strongly agree into 'Agree'
'Tend to Agree'=3
'Strongly Agree'=4

These values will be combined into a single new variable created for each of survey questions we are interested in (listed above).

In the new variable 'Disagree'=1 and 'Agree'=2.
```{r }
## Test data
ppv.tmp <- ppv

## Area_cyc_unsafe
ppv.tmp2 <- ppv.tmp %>%
  mutate(Area_cyc_unsafe_AgDis = ifelse(Area_cyc_unsafe==1 | Area_cyc_unsafe==2, 1, NA)) %>%
  mutate(Area_cyc_unsafe_AgDis = ifelse(is.na(Area_cyc_unsafe_AgDis), ifelse(Area_cyc_unsafe==3 | Area_cyc_unsafe==4, 2, Area_cyc_unsafe_AgDis), Area_cyc_unsafe_AgDis)) 
# %>%
#   select(ID, Area_cyc_unsafe, Area_cyc_unsafe_AgDis) %>%
#   sample_n(400) %>%
#   View()

## Area_cross
ppv.tmp2 <- ppv.tmp2 %>%
  mutate(Area_cross_AgDis = ifelse(Area_cross==1 | Area_cross==2, 1, NA)) %>%
  mutate(Area_cross_AgDis = ifelse(is.na(Area_cross_AgDis), ifelse(Area_cross==3 | Area_cross==4, 2, Area_cross_AgDis), Area_cross_AgDis)) 
# %>%
#   select(ID, Area_cross, Area_cross_AgDis) %>%
#   sample_n(400) %>%
#   View()


## Area_talk
ppv.tmp2 <- ppv.tmp2 %>%
  mutate(Area_talk_AgDis = ifelse(Area_talk==1 | Area_talk==2, 1, NA)) %>%
  mutate(Area_talk_AgDis = ifelse(is.na(Area_talk_AgDis), ifelse(Area_talk==3 | Area_talk==4, 2, Area_talk_AgDis), Area_talk_AgDis)) 
# %>%
# select(ID, Area_talk, Area_talk_AgDis) %>%
#   sample_n(400) %>%
#   View()


## Area_8yo_walk
ppv.tmp2 <- ppv.tmp2 %>%
  mutate(Area_8yo_walk_AgDis = ifelse(Area_8yo_walk==1 | Area_8yo_walk==2, 1, NA)) %>%
  mutate(Area_8yo_walk_AgDis = ifelse(is.na(Area_8yo_walk_AgDis), ifelse(Area_8yo_walk==3 | Area_8yo_walk==4, 2, Area_8yo_walk_AgDis), Area_8yo_walk_AgDis)) 
# %>%
#   select(ID, Area_8yo_walk, Area_8yo_walk_AgDis) %>%
#   sample_n(400) %>%
#   View()


## Area_stop
ppv.tmp2 <- ppv.tmp2 %>%
  mutate(Area_stop_AgDis = ifelse(Area_stop==1 | Area_stop==2, 1, NA)) %>%
  mutate(Area_stop_AgDis = ifelse(is.na(Area_stop_AgDis), ifelse(Area_stop==3 | Area_stop==4, 2, Area_stop_AgDis), Area_stop_AgDis)) 
# %>%
#   select(ID, Area_stop, Area_stop_AgDis) %>%
#   sample_n(400) %>%
#   View()


## Area_crime
ppv.tmp2 <- ppv.tmp2 %>%
  mutate(Area_crime_AgDis = ifelse(Area_crime==1 | Area_crime==2, 1, NA)) %>%
  mutate(Area_crime_AgDis = ifelse(is.na(Area_crime_AgDis), ifelse(Area_crime==3 | Area_crime==4, 2, Area_crime_AgDis), Area_crime_AgDis)) 
# %>%
#   select(ID, Area_crime, Area_crime_AgDis) %>%
#   sample_n(400) %>%
#   View()


## Area_cyc_path
ppv.tmp2 <- ppv.tmp2 %>%
  mutate(Area_cyc_path_AgDis = ifelse(Area_cyc_path==1 | Area_cyc_path==2, 1, NA)) %>%
  mutate(Area_cyc_path_AgDis = ifelse(is.na(Area_cyc_path_AgDis), ifelse(Area_cyc_path==3 | Area_cyc_path==4, 2, Area_cyc_path_AgDis), Area_cyc_path_AgDis)) 
# %>%
#   select(ID, Area_cyc_path, Area_cyc_path_AgDis) %>%
#   sample_n(400) %>%
#   View()


## Area_walk_pleas
ppv.tmp2 <- ppv.tmp2 %>%
  mutate(Area_walk_pleas_AgDis = ifelse(Area_walk_pleas==1 | Area_walk_pleas==2, 1, NA)) %>%
  mutate(Area_walk_pleas_AgDis = ifelse(is.na(Area_walk_pleas_AgDis), ifelse(Area_walk_pleas==3 | Area_walk_pleas==4, 2, Area_walk_pleas_AgDis), Area_walk_pleas_AgDis)) 
# %>%
#   select(ID, Area_walk_pleas, Area_walk_pleas_AgDis) %>%
#   sample_n(400) %>%
#   View()


## Area_shade
ppv.tmp2 <- ppv.tmp2 %>%
  mutate(Area_shade_AgDis = ifelse(Area_shade==1 | Area_shade==2, 1, NA)) %>%
  mutate(Area_shade_AgDis = ifelse(is.na(Area_shade_AgDis), ifelse(Area_shade==3 | Area_shade==4, 2, Area_shade_AgDis), Area_shade_AgDis)) 
# %>%
#   select(ID, Area_shade, Area_shade_AgDis) %>%
#   sample_n(400) %>%
#   View()
# 


## Area_airpoll
ppv.tmp2 <- ppv.tmp2 %>%
  mutate(Area_airpoll_AgDis = ifelse(Area_airpoll==1 | Area_airpoll==2, 1, NA)) %>%
  mutate(Area_airpoll_AgDis = ifelse(is.na(Area_airpoll_AgDis), ifelse(Area_airpoll==3 | Area_airpoll==4, 2, Area_airpoll_AgDis), Area_airpoll_AgDis)) 
# %>%
#   select(ID, Area_airpoll, Area_airpoll_AgDis) %>%
#   sample_n(400) %>%
#   View()


## Area_cyc_pleas
ppv.tmp2 <- ppv.tmp2 %>%
  mutate(Area_cyc_pleas_AgDis = ifelse(Area_cyc_pleas==1 | Area_cyc_pleas==2, 1, NA)) %>%
  mutate(Area_cyc_pleas_AgDis = ifelse(is.na(Area_cyc_pleas_AgDis), ifelse(Area_cyc_pleas==3 | Area_cyc_pleas==4, 2, Area_cyc_pleas_AgDis), Area_cyc_pleas_AgDis)) 
# %>%
#   select(ID, Area_cyc_pleas, Area_cyc_pleas_AgDis) %>%
#   sample_n(400) %>%
#   View()


## Area_8yo_cyc
ppv.tmp2 <- ppv.tmp2 %>%
  mutate(Area_8yo_cyc_AgDis = ifelse(Area_8yo_cyc==1 | Area_8yo_cyc==2, 1, NA)) %>%
  mutate(Area_8yo_cyc_AgDis = ifelse(is.na(Area_8yo_cyc_AgDis), ifelse(Area_8yo_cyc==3 | Area_8yo_cyc==4, 2, Area_8yo_cyc_AgDis), Area_8yo_cyc_AgDis)) 
# %>%
#   select(ID, Area_8yo_cyc, Area_8yo_cyc_AgDis) %>%
#   sample_n(400) %>%
#   View()


## Area_facilities
ppv.tmp2 <- ppv.tmp2 %>%
  mutate(Area_facilities_AgDis = ifelse(Area_facilities==1 | Area_facilities==2, 1, NA)) %>%
  mutate(Area_facilities_AgDis = ifelse(is.na(Area_facilities_AgDis), ifelse(Area_facilities==3 | Area_facilities==4, 2, Area_facilities_AgDis), Area_facilities_AgDis)) 
# %>%
#   select(ID, Area_facilities, Area_facilities_AgDis) %>%
#   sample_n(400) %>%
#   View()
# 


## Area_walk_unsafe
ppv.tmp2 <- ppv.tmp2 %>%
  mutate(Area_walk_unsafe_AgDis = ifelse(Area_walk_unsafe==1 | Area_walk_unsafe==2, 1, NA)) %>%
  mutate(Area_walk_unsafe_AgDis = ifelse(is.na(Area_walk_unsafe_AgDis), ifelse(Area_walk_unsafe==3 | Area_walk_unsafe==4, 2, Area_walk_unsafe_AgDis), Area_walk_unsafe_AgDis)) 
# %>%
#   select(ID, Area_walk_unsafe, Area_walk_unsafe_AgDis) %>%
#   sample_n(400) %>%
#   View()


## Area_pavements
ppv.tmp2 <- ppv.tmp2 %>%
  mutate(Area_pavements_AgDis = ifelse(Area_pavements==1 | Area_pavements==2, 1, NA)) %>%
  mutate(Area_pavements_AgDis = ifelse(is.na(Area_pavements_AgDis), ifelse(Area_pavements==3 | Area_pavements==4, 2, Area_pavements_AgDis), Area_pavements_AgDis)) 
# %>%
#   select(ID, Area_pavements, Area_pavements_AgDis) %>%
#   sample_n(400) %>%
#   View()

```

```{r Check new variables}
ppv.tmp2 %>%
  glimpse()

## Set
ppv <- ppv.tmp2

## Clean up
rm(ppv.tmp, ppv.tmp2)
```



```{r Function for variable generation, include=F, eval=F}
## Explore function for this at later stage
ppv.AgDis <- ppv.tmp %>%
  select(ID, starts_with("Area"), -Areacomments)


mut_AgDis <- function(DF) {
  require(dplyr)
  for (i in 2:ncol(DF)) { 
    paste0(DF[i],"_AgDis")
  }

}

mut_AgDis(ppv.AgDis)
```







