---
title: "Personal Information"
author: "Mark Ruddy"
date: "29 July 2016"
output: html_document
---

Load libraries and data
```{r Setup, include=FALSE}
library(dplyr)
library(tidyr)
library(stringr)
library(knitr)
library(ggplot2)
# setwd("/Users/markruddy/Documents/Personal/Work/DataScience/PeopleAndPlaces/")
# load("./Data/pp.data.RData")
```


## Tasks

1. *Gender*: Create dummy gender variable with male, female, other(missing).
2. *Age*: Check and change to numerical variable, or missing (for refused or invalid) and also create new variable for different age categories, suggest under 25/25-34,35-44,45-54,55-64,65+
3. *Ethnicity*: Create new variable that is white/BAME or missing/refused
4. *Disability*: Change 'prefer not to say' to user-missing
5. *Disability_oth*: Check and recode if necessary before creating new single variable


1. *Gender*: Create dummy gender variable with male, female, other(missing).

New variable *Gender_summary*. *Gender* will map to *Gender_summary* as follows:

|Option|*Gender*|*Gender_summary*|
|:----|:-------|:---------------|
|Male |1       |1               |
|Female|2     | 2               |
|Other|3      | 3               |
|Prefer not to say|4  |3        |
|Missing  | -99   |3            |
|NA       |NA     |3            |

```{r Creat Gender_summary variable}
## test data
ppv.tmp <- ppv

ppv.tmp2 <- ppv.tmp %>%
  mutate(Gender_summary = ifelse(Gender==1,1,NA)) %>%
  mutate(Gender_summary = ifelse(is.na(Gender_summary),ifelse(Gender==2, 2, Gender_summary), Gender_summary)) %>%
  mutate(Gender_summary = ifelse(is.na(Gender_summary), 3, Gender_summary)) 
# %>%
#     select(ID, Gender, Gender_summary) %>%
#     View()

## Check
ppv.tmp2 %>%
  select(ID, Gender, Gender_summary) %>%
    View()

glimpse(ppv.tmp2)
## OK

## Set changes
ppv <- ppv.tmp2

## Clean up
rm(ppv.tmp, ppv.tmp2)
```




## 2. *Age*: 

### Check and change to numerical variable, or missing (for refused or invalid) 

```{r Change Age to numeric}

## Change to numeric
ppv.tmp2 <- ppv.tmp %>%
  mutate(Age_num = as.numeric(Age))

## Check
ppv.tmp2 %>%
  select(ID, Age, Age_num) %>%
  arrange(desc(Age))
## OK

## Set change
ppv.tmp <- ppv.tmp %>%
  mutate(Age = as.numeric(Age))

ppv.tmp %>%
  select(ID, Age) %>%
  arrange(desc(Age))
## OK

## Set
ppv <- ppv.tmp

## Clean up
rm(ppv.tmp, ppv.tmp2)
```

Check age entries that are less than 10 or greater than 80.
These will be flagged in a new variable *Age_query*.
```{r Check Age values}
## Test data
ppv.tmp <- ppv

## List and count ages
ppv.tmp %>%
  select(Age) %>%
  group_by(Age) %>%
  summarise(count=n()) %>%
  View()

## Query low and high ages
ppv.tmp %>%
  select(ID, Age, Household) %>%
  filter(Age<10 | Age>80) %>%
  arrange(desc(Age)) %>%
  kable()
```
There are 35 responses with Age less than 10 or greater than 80.

Create Age_query variable. Where 1 = ages less than 10 and greater than 80.
```{r Create Age_query}
## Create Age_query
ppv.tmp <- ppv.tmp %>%
  mutate(Age_query = ifelse(Age<10 | Age>80, 1, NA))

## Check
ppv.tmp %>%
  select(ID, Age, Age_query) %>%
  View()
## OK

## Set changes
ppv <- ppv.tmp

## Clean up
rm(ppv.tmp)
```

### Create new variable for different age categories, suggest under 25, 25-34, 35-44, 45-54, 55-64, 65+


new *Age_cat* variable sumamry:

| Age-range | Age_cat value |
|:----------|:--------------|
|under 25   |1              |
|25-34      |2              |
|35-44      |3              |
|45-54      |4              |
|55-64      |5              |
|65+        |6              |

```{r Age categories}
## Test data
ppv.tmp <- ppv

## Create variable
ppv.tmp2 <- ppv.tmp %>%
  mutate(Age_cat = ifelse(Age<25, 1, NA)) %>%
  mutate(Age_cat = ifelse(is.na(Age_cat), ifelse(Age>=25 & Age<=34, 2, Age_cat), Age_cat)) %>%
  mutate(Age_cat = ifelse(is.na(Age_cat), ifelse(Age>=35 & Age<=44, 3, Age_cat), Age_cat)) %>%
  mutate(Age_cat = ifelse(is.na(Age_cat), ifelse(Age>=45 & Age<=54, 4, Age_cat), Age_cat)) %>%
  mutate(Age_cat = ifelse(is.na(Age_cat), ifelse(Age>=55 & Age<=64, 5, Age_cat), Age_cat)) %>%
  mutate(Age_cat = ifelse(is.na(Age_cat), ifelse(Age>=65, 6, Age_cat), Age_cat)) 
# %>%
#   select(ID, Age, Age_cat) %>%
#   sample_n(500) %>%
#   View()

## Check
ppv.tmp2 %>%
  select(ID, Age, Age_cat) %>%
  sample_n(500) %>%
  View()
## OK

## Set changes
ppv <- ppv.tmp2

## Clean up
rm(ppv.tmp, ppv.tmp2)
```

## 3. *Ethnicity*: Create new variable that is white/BAME or missing/refused

New *Ethnicity_cat* variable comparison with *Ethnicity*:

| *Ethnicity* | *Ethnicity* value | *Ethnicity_cat* value |
|:---------------|:--------------|:----------------------|
|White                |1           |1              |
|Mixed ethnic group    |2          |1               |
|Asian or Asian British|3         |1               |
|Black or Black British|4         |1              |
|Any other ethnic group|5          |1              |
|Prefer not to say    |6           |2               |
|*Missing*            |-99         |2               |
|NA                   |NA         |2               |


```{r Scope Ethnicity}
## Test data
ppv.tmp <- ppv

## List and count ages
ppv.tmp %>%
  select(Ethnicity) %>%
  group_by(Ethnicity) %>%
  summarise(count=n()) %>%
  View()

```


```{r Create Ethnicity_cat variable}
ppv.tmp2 <- ppv.tmp %>%
  mutate(Ethnicity_cat = ifelse(Ethnicity>=1 & Ethnicity<=5, 1, NA)) %>%
  mutate(Ethnicity_cat = ifelse(is.na(Ethnicity_cat), ifelse(Ethnicity==6 | Ethnicity==-99, 2, Ethnicity_cat), Ethnicity_cat)) %>%
  mutate(Ethnicity_cat = ifelse(is.na(Ethnicity_cat), 2, Ethnicity_cat))
# %>%
#   View()

## Check
ppv.tmp2 %>%
  select(ID, Ethnicity, Ethnicity_cat) %>%
  sample_n(450) %>%
  View()

## Set changes
ppv <- ppv.tmp2

## Clean up
rm(ppv.tmp2, ppv.tmp)
```


## 4. *Disability*: Change 'prefer not to say' to user-missing

Change *Disability*=6 to -99.
```{r Change Disability value}
##Test data
ppv.tmp <- ppv

## Scope dataset
ppv.tmp %>%
  select(Disability) %>%
  filter(Disability==6) %>%
  View()
## There are 136 Disability=6 entries

## Change Disability=6 to -99
ppv.tmp2 <- ppv.tmp %>%
  mutate(Disability = ifelse(Disability==6, -99, Disability))
# %>%
#   select(ID, Disability, Disability_new) %>%
#   View()

## Check
ppv.tmp2 %>%
  select(ID, Disability) %>%
  View()
## OK

## Set changes
ppv <- ppv.tmp2

## Clean up
rm(ppv.tmp2, ppv.tmp)
```


## 5. *Disability_oth*: Check and recode if necessary before creating new single variable

```{r Check Disability_oth values}

ppv.tmp <- ppv

ppv.dis <- ppv.tmp %>%
  select(ID, Disability_othtxt) %>%
  mutate(Disability_reassign = 0) %>%
  na.omit(Disability_othtxt) %>%
  arrange(ID)

```
There are 89 entries of different 'other' disabilities.

Attempt to assign *Disability_othtxt* entries to one of seven other *Disability_?* variables:

* Disability_mob
* Disability_age
* Disability_VI
* Disability_HI
* Disability_LD
* Disability_MH
* Disability_LTI

A new variable *Disability_reassign* will be created to contain a text strings of the name of the other *Disability_?* variable that seems to be appropriate. This will allow the choices to be vetted and confirmed.

Summary of major disability groups from *Disability_othtxt*
```{r Interrogate and summarise}

ppv.dis %>%
  mutate(Disabil_grp = ifelse(grepl("Autis|autism", Disability_othtxt), "Autism", NA)) %>%
  mutate(Disabil_grp = ifelse(grepl("Asthma|asthma|breath|Breath", Disability_othtxt), "Asthma", Disabil_grp)) %>%
  mutate(Disabil_grp = ifelse(grepl("Heart|heart", Disability_othtxt), "Heart", Disabil_grp)) %>%
  mutate(Disabil_grp = ifelse(grepl("Diabet|diabet", Disability_othtxt), "Diabetes", Disabil_grp)) %>%
  mutate(Disabil_grp = ifelse(grepl("Arth|arth", Disability_othtxt), "Arthritis", Disabil_grp)) %>%
  group_by(Disabil_grp) %>%
  na.omit(Disabil_grp) %>%
  summarise(count=n()) %>%
  arrange(desc(count)) %>%
  kable(title="Some major groupings 'other' disability conditions", col.names=c("Disability group","count"))
```
Other text descriptions will be reassigned to *Disability_reassign* on an ad hoc basis as set out in the code below through listing the response IDs involved.

```{r Assign Disability_othtext to Disability_reassign, include=FALSE}
## Working dataset
ppv.dis.re <- ppv.dis

## Query Disability_othtxt entries
ppv.dis.re %>%
  # filter(grepl("Autis|autism", Disability_othtxt)) 
  # slice(81:89) %>%
  View()

## Assign entries to existing options
ppv.dis.re <- ppv.dis.re %>%
  mutate(Disability_reassign = ifelse(grepl("Autis|autism", Disability_othtxt), "Disability_MH", NA)) %>%
  mutate(Disability_reassign = ifelse(grepl("Asthma|asthma|breath|Breath", Disability_othtxt), "Disability_LTI", Disability_reassign)) %>%
  mutate(Disability_reassign = ifelse(grepl("Heart|heart", Disability_othtxt), "Disability_LTI", Disability_reassign)) %>%
  mutate(Disability_reassign = ifelse(grepl("Diabet|diabet", Disability_othtxt), "Disability_LTI", Disability_reassign)) %>%
  mutate(Disability_reassign = ifelse(grepl("Arth|arth", Disability_othtxt), "Disability_mob", Disability_reassign)) %>%
  mutate(Disability_reassign = ifelse(ID %in% c(32,422,829,882,1251,1304,1394,1417,1550,1645,1645,1654,2366,2371,2396,2599,2774,3964,4063,4083,4136,4222,4765,4961,5088), "Disability_LTI", Disability_reassign)) %>%
  mutate(Disability_reassign = ifelse(ID %in% c(634,718,862,1140,1606,2026,2317,4025,4650,4736,4774,5111,5317), "Disability_mob", Disability_reassign)) %>%
  mutate(Disability_reassign = ifelse(ID %in% c(1033,4561), "Disability_age", Disability_reassign)) %>%
  mutate(Disability_reassign = ifelse(ID %in% c(1223,4626,4322), "Disability_MH", Disability_reassign))

## Check missing
ppv.dis.re %>%
  # filter(!is.na(Disability_reassign)) %>%
  filter(is.na(Disability_reassign) & !is.na(Disability_othtxt)) %>%
  View()
## All but one filled
```
All but one entry was reassigned: response ID 2703 *Disability_othtxt* was '...'.



```{r Add Disability_reassign to dataset}
## Make changes
ppv.tmp2 <- ppv.dis.re %>%
  select(-Disability_othtxt) %>%
  right_join(ppv.tmp, "ID")

## Check changes
ppv.tmp2 %>%
  select(ID, Disability_reassign, Disability, Disability_othtxt) %>%
  View()

glimpse(ppv.tmp2)
## OK

## Set changes
ppv <- ppv.tmp2

## Clean up
rm(ppv.tmp2, ppv.tmp, ppv.dis, ppv.dis.re)
```

Employing strings as variables in *Disability_reassign* is not consistent with other variables in the dataset. The strings will here be changed to the following numerical values:

|Value      |Label      |
|:----------|:-----------|
|1          |Disability_MH|
|2          |Disability_LTI|
|3          |Disability_mob|
|4          |Disability_age|
|5          |Disability_othtxt|

```{r Disability_reassign with Value:Label pairs}

ppv.tmp <- ppv

ppv.tmp2 <- ppv.tmp %>%
  mutate(Disability_reassign = ifelse(Disability_reassign=="Disability_MH",1,Disability_reassign)) %>%
  mutate(Disability_reassign = ifelse(Disability_reassign=="Disability_LTI",2,Disability_reassign)) %>%
  mutate(Disability_reassign = ifelse(Disability_reassign=="Disability_mob",3,Disability_reassign)) %>%
  mutate(Disability_reassign = ifelse(Disability_reassign=="Disability_age",4,Disability_reassign)) %>%
  mutate(Disability_reassign = ifelse(ID==2703,5,Disability_reassign))

## Check
ppv.tmp2 %>%
  select(ID, Disability_reassign) %>%
  group_by(Disability_reassign) %>%
  summarise(count=n())
## OK

## Set changes
ppv <- ppv.tmp2

rm(ppv.tmp, ppv.tmp2)

```





