---
title: "SPSS value labels"
author: "Mark Ruddy"
date: "2 August 2016"
output: html_document
---

## Task

Compile list of variable names, their values and value labels.

Then format them in the manner:

VARNAME1 ADD VALUE LABELS 1 "Strongly Agree or Agree" 2 "Disagree or Strongly Disagree”.
VARNAME2 ADD VALUE LABELS 1 “Male" 2 “Female”.
VARNAME3 ADD VALUE LABELS 1 “Cycle" 2 “Not cycle”.


## Compile list of variable names, their values and value labels.

### New variables created during checks

```{r Create data frame of variables values and labels.}
varname <- rep(c("Household"),6)
values <- c(1,2,3,4,5,6)
text <- c("Single adult living alone",
"Single adult living with other adults", "A couple without children", "A couple with children", "Other", "Single adult living with children")

labels.df <- data.frame(varname, values, text)

labels.df

## Clean up
# rm(varname, values, text)
```

```{r Existing variables their values and value names}
## Cancyc

varname <- rep("Cancyc",4)
values <- c(1,2,3,4)
text <- c("Yes", "No - physically unable to cycle", "No - don't know how/haven't learnt", "Refused to answer")

labels.df <- rbind(labels.df, data.frame(varname, values, text))

## Cyc12mon
varname <- rep("Cyc12mon",2)
values <- c(1,2)
text <- c("Yes", "No")

labels.df <- rbind(labels.df, data.frame(varname, values, text))

## Cycmonth
varname <- rep("Cycmonth",2)
values <- c(1,2)
text <- c("Yes", "No")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Other_ad_cyc2
varname <- rep("Other_ad_cyc2",4)
values <- c(1,2,3,4)
text <- c("Yes", "No", "Don't know", "Refused to answer")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## HH_child_cyc
varname <- rep("HH_child_cyc",4)
values <- c(1,2,3,4)
text <- c("Yes, often", "Yes, occasionally", "Never", "Refused to answer")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Commutes
varname <- rep("Commutes",3)
values <- c(1,2,3)
text <- c("Yes", "No", "Refused to answer")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Commcar
varname <- rep("Commcar",4)
values <- c(1,2,3,4)
text <- c("Always", "Usually", "Occasionally", "Rarely or never")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## CommPT
varname <- rep("CommPT",4)
values <- c(1,2,3,4)
text <- c("Always", "Usually", "Occasionally", "Rarely or never")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Commcyc
varname <- rep("Commcyc",4)
values <- c(1,2,3,4)
text <- c("Always", "Usually", "Occasionally", "Rarely or never")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Commwalk
varname <- rep("Commwalk",4)
values <- c(1,2,3,4)
text <- c("Always", "Usually", "Occasionally", "Rarely or never")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Escorts
varname <- rep("Escorts",3)
values <- c(1,2,3)
text <- c("Yes", "No", "Refused to answer")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Esccar
varname <- rep("Esccar",4)
values <- c(1,2,3,4)
text <- c("Always", "Usually", "Occasionally", "Rarely or never")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## EscPT
varname <- rep("EscPT",4)
values <- c(1,2,3,4)
text <- c("Always", "Usually", "Occasionally", "Rarely or never")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Esccyc
varname <- rep("Esccyc",4)
values <- c(1,2,3,4)
text <- c("Always", "Usually", "Occasionally", "Rarely or never")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Escwalk
varname <- rep("Escwalk",4)
values <- c(1,2,3,4)
text <- c("Always", "Usually", "Occasionally", "Rarely or never")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Food
varname <- rep("Food",3)
values <- c(1,2,3)
text <- c("Yes", "No", "Refused to answer")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Foodwalk
varname <- rep("Foodwalk",4)
values <- c(1,2,3,4)
text <- c("Always", "Usually", "Occasionally", "Rarely or never")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## FoodPT
varname <- rep("FoodPT",4)
values <- c(1,2,3,4)
text <- c("Always", "Usually", "Occasionally", "Rarely or never")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Foodcyc
varname <- rep("Foodcyc",4)
values <- c(1,2,3,4)
text <- c("Always", "Usually", "Occasionally", "Rarely or never")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Foodwalk
varname <- rep("Foodcyc",4)
values <- c(1,2,3,4)
text <- c("Always", "Usually", "Occasionally", "Rarely or never")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Foodwalk
varname <- rep("Foodcyc",4)
values <- c(1,2,3,4)
text <- c("Always", "Usually", "Occasionally", "Rarely or never")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Friends
varname <- rep("Friends",3)
values <- c(1,2,3)
text <- c("Yes", "No", "Refused to answer")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Friendcar
varname <- rep("Friendcar",4)
values <- c(1,2,3,4)
text <- c("Always", "Usually", "Occasionally", "Rarely or never")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## FriendPT
varname <- rep("FriendPT",4)
values <- c(1,2,3,4)
text <- c("Always", "Usually", "Occasionally", "Rarely or never")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Friendcyc
varname <- rep("Friendcyc",4)
values <- c(1,2,3,4)
text <- c("Always", "Usually", "Occasionally", "Rarely or never")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Friendwalk
varname <- rep("Friendwalk",4)
values <- c(1,2,3,4)
text <- c("Always", "Usually", "Occasionally", "Rarely or never")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Prefwalk
varname <- rep("Prefwalk",6)
values <- c(1,2,3,4,5,6)
text <- c("I use this and it is my preferred method of travel", "This is one of several methods I am happy to use", "I use this method but would prefer not to", "I would like to use this method, but currently don't", "I don't use this method and prefer not to", "Refused to answer")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Prefcyc
varname <- rep("Prefcyc",6)
values <- c(1,2,3,4,5,6)
text <- c("I use this and it is my preferred method of travel", "This is one of several methods I am happy to use", "I use this method but would prefer not to", "I would like to use this method, but currently don't", "I don't use this method and prefer not to", "Refused to answer")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Prefbus
varname <- rep("Prefbus",6)
values <- c(1,2,3,4,5,6)
text <- c("I use this and it is my preferred method of travel", "This is one of several methods I am happy to use", "I use this method but would prefer not to", "I would like to use this method, but currently don't", "I don't use this method and prefer not to", "Refused to answer")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Preftrain
varname <- rep("Preftrain",6)
values <- c(1,2,3,4,5,6)
text <- c("I use this and it is my preferred method of travel", "This is one of several methods I am happy to use", "I use this method but would prefer not to", "I would like to use this method, but currently don't", "I don't use this method and prefer not to", "Refused to answer")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Preftube
varname <- rep("Preftube",6)
values <- c(1,2,3,4,5,6)
text <- c("I use this and it is my preferred method of travel", "This is one of several methods I am happy to use", "I use this method but would prefer not to", "I would like to use this method, but currently don't", "I don't use this method and prefer not to", "Refused to answer")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Prefcar
varname <- rep("Prefcar",6)
values <- c(1,2,3,4,5,6)
text <- c("I use this and it is my preferred method of travel", "This is one of several methods I am happy to use", "I use this method but would prefer not to", "I would like to use this method, but currently don't", "I don't use this method and prefer not to", "Refused to answer")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Suppwalk
varname <- rep("Suppwalk",5)
values <- c(2,3,4,6,7)
text <- c("Too little", "About the right amount", "Too much", "Don't know or not applicable", "Refused to know")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Suppcyc
varname <- rep("Suppcyc",5)
values <- c(2,3,4,6,7)
text <- c("Too little", "About the right amount", "Too much", "Don't know or not applicable", "Refused to know")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Suppbus
varname <- rep("Suppbus",5)
values <- c(2,3,4,6,7)
text <- c("Too little", "About the right amount", "Too much", "Don't know or not applicable", "Refused to know")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Supptrain
varname <- rep("Supptrain",5)
values <- c(2,3,4,6,7)
text <- c("Too little", "About the right amount", "Too much", "Don't know or not applicable", "Refused to know")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Supptube
varname <- rep("Supptube",5)
values <- c(2,3,4,6,7)
text <- c("Too little", "About the right amount", "Too much", "Don't know or not applicable", "Refused to know")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Suppcar
varname <- rep("Suppcar",5)
values <- c(2,3,4,6,7)
text <- c("Too little", "About the right amount", "Too much", "Don't know or not applicable", "Refused to know")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Area_cyc_unsafe
varname <- rep("Area_cyc_unsafe",6)
values <- c(1,2,3,4,5,8)
text <- c("Strongly Disagree", "Tend to Disagree", "Tend to Agree", "Strongly Agree", "Neither Agree nor Disagree", "Refused to Answer")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Area_cross
varname <- rep("Area_cross",6)
values <- c(1,2,3,4,5,8)
text <- c("Strongly Disagree", "Tend to Disagree", "Tend to Agree", "Strongly Agree", "Neither Agree nor Disagree", "Refused to Answer")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Area_talk
varname <- rep("Area_talk",6)
values <- c(1,2,3,4,5,8)
text <- c("Strongly Disagree", "Tend to Disagree", "Tend to Agree", "Strongly Agree", "Neither Agree nor Disagree", "Refused to Answer")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Area_8yo_walk
varname <- rep("Area_8yo_walk",6)
values <- c(1,2,3,4,5,8)
text <- c("Strongly Disagree", "Tend to Disagree", "Tend to Agree", "Strongly Agree", "Neither Agree nor Disagree", "Refused to Answer")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Area_stop
varname <- rep("Area_stop",6)
values <- c(1,2,3,4,5,8)
text <- c("Strongly Disagree", "Tend to Disagree", "Tend to Agree", "Strongly Agree", "Neither Agree nor Disagree", "Refused to Answer")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Area_crime
varname <- rep("Area_crime",6)
values <- c(1,2,3,4,5,8)
text <- c("Strongly Disagree", "Tend to Disagree", "Tend to Agree", "Strongly Agree", "Neither Agree nor Disagree", "Refused to Answer")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Area_cyc_path
varname <- rep("Area_cyc_path",6)
values <- c(1,2,3,4,5,8)
text <- c("Strongly Disagree", "Tend to Disagree", "Tend to Agree", "Strongly Agree", "Neither Agree nor Disagree", "Refused to Answer")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Area_walk_pleas
varname <- rep("Area_walk_pleas",6)
values <- c(1,2,3,4,5,8)
text <- c("Strongly Disagree", "Tend to Disagree", "Tend to Agree", "Strongly Agree", "Neither Agree nor Disagree", "Refused to Answer")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Area_shade
varname <- rep("Area_shade",6)
values <- c(1,2,3,4,5,8)
text <- c("Strongly Disagree", "Tend to Disagree", "Tend to Agree", "Strongly Agree", "Neither Agree nor Disagree", "Refused to Answer")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Area_airpoll
varname <- rep("Area_airpoll",6)
values <- c(1,2,3,4,5,8)
text <- c("Strongly Disagree", "Tend to Disagree", "Tend to Agree", "Strongly Agree", "Neither Agree nor Disagree", "Refused to Answer")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Area_cyc_pleas
varname <- rep("Area_cyc_pleas",6)
values <- c(1,2,3,4,5,8)
text <- c("Strongly Disagree", "Tend to Disagree", "Tend to Agree", "Strongly Agree", "Neither Agree nor Disagree", "Refused to Answer")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Area_8yo_cyc
varname <- rep("Area_8yo_cyc",6)
values <- c(1,2,3,4,5,8)
text <- c("Strongly Disagree", "Tend to Disagree", "Tend to Agree", "Strongly Agree", "Neither Agree nor Disagree", "Refused to Answer")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Area_facilities
varname <- rep("Area_facilities",6)
values <- c(1,2,3,4,5,8)
text <- c("Strongly Disagree", "Tend to Disagree", "Tend to Agree", "Strongly Agree", "Neither Agree nor Disagree", "Refused to Answer")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Area_walk_unsafe
varname <- rep("Area_walk_unsafe",6)
values <- c(1,2,3,4,5,8)
text <- c("Strongly Disagree", "Tend to Disagree", "Tend to Agree", "Strongly Agree", "Neither Agree nor Disagree", "Refused to Answer")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Area_pavements
varname <- rep("Area_pavements",6)
values <- c(1,2,3,4,5,8)
text <- c("Strongly Disagree", "Tend to Disagree", "Tend to Agree", "Strongly Agree", "Neither Agree nor Disagree", "Refused to Answer")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Noise_home
varname <- rep("Noise_home",6)
values <- c(1,2,3,4,5,8)
text <- c("Extremely", "Very", "Slightly", "Not at all", "Moderately", "Refused to answer")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Noise_area
varname <- rep("Noise_area",6)
values <- c(1,2,3,4,5,8)
text <- c("Extremely", "Very", "Slightly", "Not at all", "Moderately", "Refused to answer")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Gender
varname <- rep("Gender",4)
values <- c(1,2,3,4)
text <- c("Male", "Female", "Other, please state", "Prefer not to say")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Ethnicity
varname <- rep("Ethnicity",6)
values <- c(1,2,3,4,5,6)
text <- c("White", "Mixed ethnic group", "Asian or Asian British", "Black or Black British", "Any other ethnic group", "Prefer not to say")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Disability
varname <- rep("Disability",3)
values <- c(4,5,6)
text <- c("Yes", "No", "Prefer not to say")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## HH_income
varname <- rep("HH_income",5)
values <- c(1,2,3,4,5)
text <- c("Under £20,000", "£20,000-£49,999", "£50,000-74,999", "Don't know / prefer not to say", "£75,000 or more")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Consent_nxt_yr
varname <- rep("Consent_nxt_yr",2)
values <- c(1,2)
text <- c("Yes", "No")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## SurveyGp
varname <- rep("SurveyGp",4)
values <- c(1,2,3,4)
text <- c("Leaflet", "Oyster database", "Cyclist database", "Unknown (Oyster or cyclist database, unmatched")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Borough
varname <- rep("Borough",4)
values <- c(1,2,3,4)
text <- c("Enfield", "Kingston", "Waltham Forest", "Control")

labels.df <- rbind(labels.df, data.frame(varname, values, text))

# labels.df %>% View()

# varname <- rep()
# values <- c()
# text <- c()

```



```{r New variables their values and value names}


## HH_text_code
varname <- rep(c("HH_text_code"),6)
values <- c(1,2,3,4,5,6)
text <- c("Single adult living alone",
"Single adult living with other adults", "A couple without children", "A couple with children", "Other", "Single adult living with children")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## HH_text_new
varname <- rep(c("HH_text_new"),6)
values <- c(1,2,3,4,5,6)
text <- c("Single adult living alone",
"Single adult living with other adults", "A couple without children", "A couple with children", "Other", "Single adult living with children")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## HH_num_age_dis
varname <- rep(c("HH_num_age_dis"),2)
values <- c(1,2)
text <- c("Real disagreement", "Missing value disagreement")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## HH_carvan_sum
varname <- rep(c("HH_carvan_sum"),3)
values <- c(0,1,2)
text <- c("No car", "One car", "Two or more cars")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Area_cyc_unsafe_AgDis
varname <- rep(c("Area_cyc_unsafe_AgDis"),2)
values <- c(1,2)
text <- c("Disagree", "Agree")

labels.df <- rbind(labels.df, data.frame(varname, values, text))

## Area_cross_AgDis
varname <- rep(c("Area_cross_AgDis"),2)
values <- c(1,2)
text <- c("Disagree", "Agree")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Area_talk_AgDis
varname <- rep(c("Area_talk_AgDis"),2)
values <- c(1,2)
text <- c("Disagree", "Agree")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Area_8yo_walk_AgDis
varname <- rep(c("Area_8yo_walk_AgDis"),2)
values <- c(1,2)
text <- c("Disagree", "Agree")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Area_stop_AgDis
varname <- rep(c("Area_stop_AgDis"),2)
values <- c(1,2)
text <- c("Disagree", "Agree")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Area_crime_AgDis
varname <- rep(c("Area_crime_AgDis"),2)
values <- c(1,2)
text <- c("Disagree", "Agree")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Area_cyc_path_AgDis
varname <- rep(c("Area_cyc_path_AgDis"),2)
values <- c(1,2)
text <- c("Disagree", "Agree")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Area_walk_pleas_AgDis
varname <- rep(c("Area_walk_pleas_AgDis"),2)
values <- c(1,2)
text <- c("Disagree", "Agree")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Area_shade_AgDis
varname <- rep(c("Area_shade_AgDis"),2)
values <- c(1,2)
text <- c("Disagree", "Agree")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Area_airpoll_AgDis
varname <- rep(c("Area_airpoll_AgDis"),2)
values <- c(1,2)
text <- c("Disagree", "Agree")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Area_cyc_pleas_AgDis
varname <- rep(c("Area_cyc_pleas_AgDis"),2)
values <- c(1,2)
text <- c("Disagree", "Agree")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Area_8yo_cyc_AgDis
varname <- rep(c("Area_8yo_cyc_AgDis"),2)
values <- c(1,2)
text <- c("Disagree", "Agree")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Area_facilities_AgDis
varname <- rep(c("Area_facilities_AgDis"),2)
values <- c(1,2)
text <- c("Disagree", "Agree")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Area_walk_unsafe_AgDis
varname <- rep(c("Area_walk_unsafe_AgDis"),2)
values <- c(1,2)
text <- c("Disagree", "Agree")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Area_pavements_AgDis
varname <- rep(c("Area_pavements_AgDis"),2)
values <- c(1,2)
text <- c("Disagree", "Agree")

labels.df <- rbind(labels.df, data.frame(varname, values, text))



## Gender_summary
varname <- rep(c("Gender_summary"),3)
values <- c(1,2,3)
text <- c("Male", "Female", "Other / Missing")

labels.df <- rbind(labels.df, data.frame(varname, values, text))



## Age_cat
varname <- rep(c("Age_cat"),6)
values <- c(1,2,3,4,5,6)
text <- c("Under 25", "25-34", "35-44", "45-54", "55-64", "65+")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Ethnicity_cat
varname <- rep(c("Ethnicity_cat"),2)
values <- c(1,2)
text <- c("White / BAME", "Other / Missing")

labels.df <- rbind(labels.df, data.frame(varname, values, text))



## Disability_reassign
varname <- rep("Disability_reassign", 5)
values <- c(1,2,3,4,5)
text <- c("Disability_MH", "Disability_LTI", "Disability_mob", "Disability_age", "Disability_othtxt")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Phyact_qry
varname <- rep("Phyact_qry", 2)
values <- c(1,2)
text <- c("Phyact < diary", "Phyact > diary")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Emp_paid
varname <- rep("Emp_paid", 2)
values <- c(1,2)
text <- c("Paid employment", "Not in paid employment")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


## Emp_FT
varname <- rep("Emp_FT", 2)
values <- c(1,2)
text <- c("Paid full-time employment", "Not in paid full-time employment")

labels.df <- rbind(labels.df, data.frame(varname, values, text))


# labels.df %>% View()
# rm(varname, values, text)
```

## Code up SPSS syntax

For each variable, compile statements in the format:

VARNAME1 ADD VALUE LABELS 1 "Strongly Agree or Agree" 2 "Disagree or Strongly Disagree”.
VARNAME2 ADD VALUE LABELS 1 “Male" 2 “Female”.
VARNAME3 ADD VALUE LABELS 1 “Cycle" 2 “Not cycle”.

```{r }

l.df <- labels.df

## Function to write SPSS syntax from labels data.frame
SPSS.labels <- function(DF) {
  syn.all <- NULL;
  names <- as.character(unique(DF[[1]]));
  for (i in 1:length(names)) {
    syn.line <- NULL;
    x <- names[i];
    DFgrp <- DF[which(DF[1]==x),];
    for (j in 1:dim(DFgrp)[1]) {
      syn.frag <- paste0(" ", DFgrp[j,2], " \"", DFgrp[j,3],"\"", sep="");
      syn.line <- paste0(syn.line, syn.frag, sep="");
      }
    syn.all <- paste0(syn.all, paste0(names[i], " ADD VALUE LABELS", syn.line, ".\n", sep=""), sep="");
  }
  return(syn.all);
}


spss.syntax <- SPSS.labels(l.df)
setwd("/Users/markruddy/Documents/Personal/Work/DataScience/PeopleAndPlaces/")
write(spss.syntax, file="./Working/Outputs/Code/spss_var_value_syntax.txt")

```

















