---
title: "Cycleweek consistency checks."
author: "Mark Ruddy"
date: "17 June 2016"
output: html_document
---

```{r Load packages, include=FALSE}
library(dplyr)
library(tidyr)
library(stringr)
```

```{r Load data}
## Load data
load("~/Documents/Personal/Work/DataScience/PeopleAndPlaces/Data/Original/pp.data.RData")
dim(pp)
```

# Cycleweek Tasks

1. Check that Cycleweek are 0 or 1 only if other trip indicator variables agree (could be zero by default in survey).
2. Create new variable flagging up inconsistencies.

Need to compare Cycleweek with other indicators of cycling in last week:

* MinsCycling
* Diary1cyc, Diary2cyc, Diary3cyc, Diary4cyc, Diary5cyc, Diary6cyc, Diary7cyc
* Cyc1min, Cyc2min, Cyc3min, Cyc4min, Cyc5min, Cyc6min, Cyc7min


## Cycleweek entries check.

Are any Cycleweek entries entered as 'refused', other strings or NA?

```{r Cycleweek entries}
## Check 'refused' responses - should just be 0 or 1
distinct(select(pp, Cycleweek))
```
There are no 'refused' entries in Cycleweek. All coded as 0 or 1.

## Cycleweek=0 and Diary?cyc=1 check.

Where Cycleweek=0, are there any cycling diary entries that are 1 (Diary?cyc=1).

```{r NO Cycleweek but YES Diary?cyc}
## Check dimensions of Cycleweek=0 & Diary?Cyc=1
ppv %>%
  filter(Cycleweek==0 | is.na(Cycleweek), Diary1cyc==1 | Diary2cyc==1 | Diary3cyc==1 | Diary4cyc==1 | Diary5cyc==1 | Diary6cyc==1 | Diary7cyc==1) %>%
  select(ID, Cycleweek, Diary1cyc, Diary2cyc, Diary3cyc, Diary4cyc, Diary5cyc, Diary6cyc, Diary7cyc) %>%
  dim()
```
There are 40 records of 'not cycling during the last week' that have an associated cycling diary entry.


Add 'noCycleweek_yesDiary' variable flagging these.
```{r Add noCycleweek_yesDiary variable}
## Add flagging variable

ppv <- ppv %>%
  mutate(noCycleweek_yesDiary = ifelse((Cycleweek==0 | is.na(Cycleweek)) & (Diary1cyc==1 | Diary2cyc==1 | Diary3cyc==1 | Diary4cyc==1 | Diary5cyc==1 | Diary6cyc==1 | Diary7cyc==1), 1, NA))

ppv %>%
  select(ID, noCycleweek_yesDiary, Cycleweek, Diary1cyc, Diary2cyc, Diary3cyc, Diary4cyc, Diary5cyc, Diary6cyc, Diary7cyc) %>%
  filter(noCycleweek_yesDiary == 1) %>%
  View()

## check that new variable records match earlier dplr Cycleweek/Diary?cyc query
## ref http://stackoverflow.com/a/21629102 use dply::collect to get vector from tbl_df
cw_dy <- pp %>%
  filter(Cycleweek==0, Diary1cyc==1 | Diary2cyc==1 | Diary3cyc==1 | Diary4cyc==1 | Diary5cyc==1 | Diary6cyc==1 | Diary7cyc==1) %>%
  select(ID) %>%
  collect %>% .[["ID"]]

cw_dy_var <- ppv %>%
  filter(noCycleweek_yesDiary == 1) %>%
  select(ID) %>%
  collect %>% .[["ID"]]

cw_dy %in% cw_dy_var
## yes! they match - so flag is correct

## Clean up
rm(cw_dy, cw_dy_var)
```

## Cycleweek=1 and Diary?cyc=0 check.

Where Cycleweek=1 are there any cycling diary entries that are zero (Diary?cyc=0).
```{r YES Cycleweek but NO Diary?cyc}
ppv %>%
  filter(Cycleweek==1, (Diary1cyc==0 | is.na(Diary1cyc)) & (Diary2cyc==0 | is.na(Diary2cyc)) & (Diary3cyc==0 | is.na(Diary3cyc)) & (Diary4cyc==0 | is.na(Diary4cyc)) & (Diary5cyc==0 | is.na(Diary5cyc)) & (Diary6cyc==0 | is.na(Diary6cyc)) & (Diary7cyc==0 | is.na(Diary7cyc))) %>%
  select(ID, Cycleweek, Diary1cyc, Diary2cyc, Diary3cyc, Diary4cyc, Diary5cyc, Diary6cyc, Diary7cyc) %>%
  View()
```
There are zero records where Cycleweek=1 and no cycling diary entry.


## Cycleweek and MinsCycling variable check.

What variables have been entered in MinsCycling (Cycling minutes during last week). This should just be numeric but are there NAs or letters?

```{r MinsCycling entries}
## Check for NAs
ppv %>%
  filter(is.na(MinsCycling)) %>%
  select(ID, Cycleweek, MinsCycling, Cyc1min:Cyc7min)
## No NAs
  
## Check for letters in MinsCycling responses - should just be numbers
## ref http://stackoverflow.com/a/21629102 to get character vector from tbl_df
mc <- collect(select(ppv,MinsCycling))[[1]]
pattern <- "[A-z]"
mc[str_detect(mc, pattern)]
rm(mc, pattern)
```
No NAs or letters in MinsCycling.

## Cycleweek=0 and MinsCycling>0 check.

Are there instances of Cycleweek=0 & MinsCycling>0?
```{r NO Cycleweek but YES MinsCycling}
pp %>%
  filter(Cycleweek==0 | is.na(Cycleweek), MinsCycling>0) %>%
  dim()
```
There are zero records of Cycleweek=0 with MinsCycling>0.

## Cycleweek=1 and MinsCycling=0 or NA check.

Are there instances of Cycleweek=1 but MinsCycling=0 or NA?
```{r YES Cycleweek but NO MinsCycling}
pp %>%
  filter(Cycleweek==1, MinsCycling==0 | is.na(MinsCycling)) %>%
  select(ID, Cycleweek, MinsCycling) %>%
  dim()
```
There are 22 records of Cycleweek=1 with MinsCycling=0.

Add 'yesCycleweek_noMinsCycling' variable flagging these.
```{r Add yesCycleweek_noMinsCycling variable}
## Add flagging variable
ppv <- ppv %>%
  mutate(yesCycleweek_noMinsCycling = ifelse(Cycleweek==1 & MinsCycling==0, 1, NA))

ppv %>%
  select(ID, yesCycleweek_noMinsCycling, Cycleweek, MinsCycling) %>%
  filter(yesCycleweek_noMinsCycling == 1) %>%
  View()

## check that new variable records match earlier dplyr Cycleweek/MinsCycling query
## ref http://stackoverflow.com/a/21629102 use dply::collect to get vector from tbl_df
a <- pp %>%
  filter(Cycleweek==1, MinsCycling==0) %>%
  select(ID) %>%
  collect %>% .[["ID"]]

b <- ppv %>%
  filter(yesCycleweek_noMinsCycling == 1) %>%
  select(ID) %>%
  collect %>% .[["ID"]]

a %in% b
## yes! they match - so flag is correct

## Clean up
rm(a, b)
```



## Cycleweek and Cyc?min variable check.

Check that where Cyclweek=0 doesn't have any Cyc?min entries.
```{r Cycleweek vs Cyc?min NA entries}
# Check Cycleweek==0 but Cyc?min are not NA
pp %>%
  select(ID, Cycleweek, Cyc1min:Cyc7min) %>%
  filter(Cycleweek==0, !is.na(Cyc1min) | !is.na(Cyc2min) | !is.na(Cyc3min) | !is.na(Cyc4min) | !is.na(Cyc5min) | !is.na(Cyc6min) | !is.na(Cyc7min)) %>%
  dim()
```
No Cyc?min entries with Cycleweek=0.

## Cycleweek=1 and Cyc?min=0 or NA check.

Where Cycleweek=1, are there NA entries to cycling minutes per day (Cyc?min) variables?
```{r YES Cycleweek but NO Cyc?min}
pp %>%
  select(ID, Cycleweek, Cyc1min:Cyc7min) %>%
  filter(Cycleweek==1, (is.na(Cyc1min) | Cyc1min==0) & (is.na(Cyc2min) | Cyc2min==0) & (is.na(Cyc3min) | Cyc3min==0) & (is.na(Cyc4min) | Cyc4min==0) & (is.na(Cyc5min) | Cyc5min==0) & (is.na(Cyc6min) | Cyc6min==0) & (is.na(Cyc7min) | Cyc7min==0)) %>%
  View()

## 22 records

```
There are 22 instances of Cycleweek=1 with NA or 0 values for cycling minutes.

Add 'yesCycleweek_noCycmins' variable flagging these.
```{r Add yesCycleweek_noCycmins variable}
## Add flagging variable
ppv <- ppv %>%
  mutate(yesCycleweek_noCycmins = ifelse(Cycleweek==1 & (is.na(Cyc1min) | Cyc1min==0) & (is.na(Cyc2min) | Cyc2min==0) & (is.na(Cyc3min) | Cyc3min==0) & (is.na(Cyc4min) | Cyc4min==0) & (is.na(Cyc5min) | Cyc5min==0) & (is.na(Cyc6min) | Cyc6min==0) & (is.na(Cyc7min) | Cyc7min==0), 1, NA))

ppv %>%
  select(ID, yesCycleweek_noCycmins, Cycleweek, Cyc1min) %>%
  filter(yesCycleweek_noCycmins == 1) %>%
  View()
```


? are yesCycleweek_noCycmins the same records as yesCycleweek_noMinsCycling records?
```{r Join by ID}
ppv.MinsCyc <- ppv %>%
  select(ID, yesCycleweek_noMinsCycling) %>%
  filter(yesCycleweek_noMinsCycling==1) %>%
  mutate(type = "MinsCyc")

ppv.Cycmin <- ppv %>%
  select(ID, yesCycleweek_noCycmins) %>%
  filter(yesCycleweek_noCycmins==1) %>%
  mutate(type = "Cycmins")

join <- inner_join(ppv.Cycmin, ppv.MinsCyc, by="ID")
  
View(join)

## Yes they are the same

rm(join, ppv.Cycmin, ppv.MinsCyc)
```

## Combine yesCycleweek_noMinsCycling and yesCycleweek_noCycmins

As the combination of Cycleweek=1 and Cyc?mins=0 or NA, and Cycleweek=1 and MinsCycling=0 or NA, yield the same records, they can be combined together into the same flagging variable.

```{r Combine yesCycleweek_noMinsCycling and yesCycleweek_noCycmins}
## Remove yesCycleweek_noMinsCycling and rename yesCycleweek_noCycmins variables
ppv <- ppv %>%
  select(-yesCycleweek_noMinsCycling) %>%
  rename(yesCycleweek_noMins = yesCycleweek_noCycmins)

ppv %>%
  filter(yesCycleweek_noMins==1) %>%
  select(ID, Cycleweek, yesCycleweek_noMins, MinsCycling, Cyc1min) %>%
  View()
```


## Cycleweek=0 and Cyc?min=1 check.

```{r NO Cycleweek and YES Cyc?min}
# Check Cycleweek==0 but Cyc?min... are not zero
ppv %>%
  select(ID, Cycleweek, Cyc1min:Cyc7min) %>%
  filter((Cycleweek==0 | is.na(Cycleweek)), Cyc1min>0 & Cyc2min>0 & Cyc3min>0 & Cyc4min>0 & Cyc5min>0 & Cyc6min>0 & Cyc7min>0) %>%
  dim()
```
There are zero instances of Cycleweek as zero with positive inputs for cycling minutes.











