---
title: "Walkweek consistency checks."
author: "Mark Ruddy"
date: "8th July 2016"
output: html_document
---

```{r Load packages, include=FALSE}
library(dplyr)
library(tidyr)
library(stringr)
```

```{r Load data}
load("~/Documents/Personal/Work/DataScience/PeopleAndPlaces/Data/pp.data.RData")

dim(pp)
```

# Walkweek Tasks

1. Check that Walkweek only zero if other trip indicator variables agree (could be zero by default in survey).
2. Create new variable flagging up inconsistencies.

Need to compare Walkweek with other indicators of walking in last week:

* MinsWalking
* Diary1walk, Diary2walk, Diary3walk, Diary4walk, Diary5walk, Diary6walk, Diary7walk
* Walk1min, Walk2min, Walk3min, Walk4min, Walk5min, Walk6min, Walk7min


## Walkweek 'refused' check

Are any Walkweek entries entered as 'refused', other strings or NA?

```{r Walkweek entries}
## Check 'refused' responses - should just be 0 or 1
distinct(select(pp, Walkweek))
```

There are no 'refused' entries in Walkweek. All coded as 0 or 1.


## Walkweek=0 and Diary?walk=1 check.

Where Walkweek=0, are there any walking diary entries that are 1 (Diary?walk=1).

```{r NO Walkweek but YES Diary?walk}
## Check dimensions of Walkweek=0 & Diary?walk=1
ppv %>%
  filter(Walkweek==0 | is.na(Walkweek), Diary1walk==1 | Diary2walk==1 | Diary3walk==1 | Diary4walk==1 | Diary5walk==1 | Diary6walk==1 | Diary7walk==1) %>%
  select(ID, Walkweek, Diary1walk, Diary2walk, Diary3walk, Diary4walk, Diary5walk, Diary6walk, Diary7walk) %>%
  View()
```
There are 151 records where Walkweek=0 or is NA that have an associated walking diary entry.


Add 'noWalkweek_yesDiary' variable flagging these.
```{r Add noWalkweek_yesDiary variable}
## Add flagging variable
ppv <- ppv %>%
  mutate(noWalkweek_yesDiary = ifelse((Walkweek==0 | is.na(Walkweek)) & (Diary1walk==1 | Diary2walk==1 | Diary3walk==1 | Diary4walk==1 | Diary5walk==1 | Diary6walk==1 | Diary7walk==1), 1, NA))

## Check new variable
ppv %>%
  filter(Walkweek==0, Diary1walk==1 | Diary2walk==1 | Diary3walk==1 | Diary4walk==1 | Diary5walk==1 | Diary6walk==1 | Diary7walk==1) %>%
  select(ID, noWalkweek_yesDiary, Walkweek, Diary1walk) %>%
  View()
```

## Walkweek=1 and Diary?walk=0 check.

Where Walkweek=1 are there any walking diary entries that are zero (Diary?walk=0).
```{r YES Walkweek but NO Diary?walk}
ppv %>%
  filter(Walkweek==1, (Diary1walk==0 | is.na(Diary1walk)) & (Diary2walk==0 | is.na(Diary2walk)) & (Diary3walk==0 | is.na(Diary3walk)) & (Diary4walk==0 | is.na(Diary4walk)) & (Diary5walk==0 | is.na(Diary5walk)) & (Diary6walk==0 | is.na(Diary6walk)) & (Diary7walk==0 | is.na(Diary7walk))) %>%
  select(ID, Walkweek, Diary1walk, Diary2walk, Diary3walk, Diary4walk, Diary5walk, Diary6walk, Diary7walk) %>%
  View()
```
There are zero records where Walkweek=1 and no walking diary entry.


## Walkweek and MinsWalking variable check.

What variables have been entered in MinsWalking (Walking minutes during last week). This should just be numeric but are there NAs or letters?


```{r MinsWalking entries}
## Check for NAs
ppv %>%
  filter(is.na(MinsWalking)) %>%
  select(ID, Walkweek, MinsWalking, Walk1min:Walk7min)
## No NAs

## Check for letters in MinsWalking responses - should just be numbers
## ref http://stackoverflow.com/a/21629102 to get character vector from tbl_df
mc <- collect(select(pp,MinsWalking))[[1]]

pattern <- "[A-z]"

mc[str_detect(mc, pattern)]

rm(mc, pattern)
```
No NAs or letter in MinsWalking.


## Walkweek=0 and MinsWalking>0 check.

Are there instances of Walkweek=0 & MinsWalking>0?
```{r NO Walkweek but YES MinsWalking}
pp %>%
  filter(Walkweek==0 | is.na(Walkweek), MinsWalking>0) %>%
  dim()
```
There are zero records of Walkweek=0 with MinsWalking>0.

## Walkweek=1 and MinsWalking=0 or NA check.

Are there instances of Walkweek=1 with MinsWalk=0 or NA?
```{r YES Walkweek but NO MinsWalking}
pp %>%
  filter(Walkweek==1, MinsWalking==0 | is.na(MinsWalking)) %>%
  select(ID, Walkweek, MinsWalking) %>%
  View()
```
There are 31 records of Walkweek=1 and MinsWalk=0.

Add 'yesWalkweek_noMinsWalking' variable flagging these.
```{r Add yesWalkweek_noMinsWalking variable}

ppv <- ppv %>%
  mutate(yesWalkweek_noMinsWalking = ifelse(Walkweek==1 & (MinsWalking==0 | is.na(MinsWalking)), 1, NA))

ppv %>%
  select(ID, yesWalkweek_noMinsWalking, Walkweek, MinsWalking) %>%
  filter(yesWalkweek_noMinsWalking==1) %>%
  View()

## Save data
#save.image("~/Documents/Personal/Work/DataScience/PeopleAndPlaces/Data/pp.data.RData")

```



## Walkweek and Walk?min variable check.

Check that where Walkweek=0 doesn't have any Walk?min entries.
```{r Walkweek and Walk?min NA check}
## Check Walkweek==0 but Walk?min... are not NA
ppv %>%
  select(ID, Walkweek, Walk1min:Walk7min) %>%
  filter(Walkweek==0, !is.na(Walk1min) | !is.na(Walk2min) | !is.na(Walk3min) | !is.na(Walk4min) | !is.na(Walk5min) | !is.na(Walk6min) | !is.na(Walk7min)) %>%
  View()
```
No Walk?min entries with Walkweek=0.


## Walkweek=1 and Walk?min=0 or NA check.
Where Walkweek=1, are there NA entries to walking minutes per day (Walk?min) variables?
```{r YES Walkweek but NO Walk?min}
ppv %>%
  select(ID, Walkweek, Walk1min:Walk7min) %>%
  filter(Walkweek==1, (is.na(Walk1min) | Walk1min==0) & (is.na(Walk2min) | Walk2min==0) & (is.na(Walk3min) | Walk3min==0) & (is.na(Walk4min) | Walk4min==0) & (is.na(Walk5min) | Walk5min==0) & (is.na(Walk6min) | Walk6min==0) & (is.na(Walk7min) | Walk7min==0)) %>%
  View()

## 31 records
```
There are 31 instances of Walkweek=1 with NA or 0 values for walking minutes.


Add 'yesWalkweek_noWalkmins' variable flagging these.
```{r Add yesWalkweek_noWalkmins variable}
## Add flagging variable
ppv <- ppv %>%
  mutate(yesWalkweek_noWalkmins = ifelse(Walkweek==1 & (is.na(Walk1min) | Walk1min==0) & (is.na(Walk2min) | Walk2min==0) & (is.na(Walk3min) | Walk3min==0) & (is.na(Walk4min) | Walk4min==0) & (is.na(Walk5min) | Walk5min==0) & (is.na(Walk6min) | Walk6min==0) & (is.na(Walk7min) | Walk7min==0), 1, NA))

ppv %>%
  select(ID, yesWalkweek_noWalkmins, Walkweek, Walk1min) %>%
  filter(yesWalkweek_noWalkmins == 1) %>%
  View()
```

? are yesWalkweek_noWalkmins the same records as yesWalkweek_noMinsWalking records?
```{r Join by ID}
ppv.MinsWalk <- ppv %>%
  select(ID, yesWalkweek_noMinsWalking) %>%
  filter(yesWalkweek_noMinsWalking==1) %>%
  mutate(type = "MinsWalk")

ppv.Walkmin <- ppv %>%
  select(ID, yesWalkweek_noWalkmins) %>%
  filter(yesWalkweek_noWalkmins==1) %>%
  mutate(type = "Walkmins")

join <- inner_join(ppv.Walkmin, ppv.MinsWalk, by="ID")
  
View(join)

## Yes they are the same

#rm(join, ppv.Walkmin, ppv.MinsWalk)
```

## Combine yesWalkweek_noMinsWalking and yesWalkweek_noWalkmins

As the combination of Walkweek=1 and Walk?mins=0 or NA, and Walkweek=1 and MinsWalking=0 or NA, yield the same records, they can be combined together into the same flagging variable.

```{r Combine yesWalkweek_noMinsWalking and yesWalkweek_noWalkmins}
## Remove yesWalkweek_noMinsWalking and rename yesWalkweek_noWalkmins variables
ppv <- ppv %>%
  select(-yesWalkweek_noMinsWalking) %>%
  rename(yesWalkweek_noMins = yesWalkweek_noWalkmins)

ppv %>%
  filter(yesWalkweek_noMins==1) %>%
  select(ID, Cycleweek, yesWalkweek_noMins, MinsWalking, Walk1min) %>%
  View()
```

## Walkweek=0 and Walk?min=1 check.
```{r NO Walkweek and YES Walk?min}
# Check Walkweek==0 but Walk?min... are not zero
ppv %>%
  select(ID, Walkweek, Walk1min:Walk7min) %>%
  filter((Walkweek==0 | is.na(Walkweek)), Walk1min>0 & Walk2min>0 & Walk3min>0 & Walk4min>0 & Walk5min>0 & Walk6min>0 & Walk7min>0) %>%
  dim()
```




















