---
title: "Carweek consistency checks."
author: "Mark Ruddy"
date: "8th July 2016"
output: html_document
---

```{r Load packages, include=FALSE}
library(dplyr)
library(tidyr)
library(stringr)
```

```{r Load data}
load("~/Documents/Personal/Work/DataScience/PeopleAndPlaces/Data/pp.data.RData")

dim(pp)
```

# Carweek Tasks

1. Check that Carweek only zero if other trip indicator variables agree (could be zero by default in survey).
2. Create new variable flagging up inconsistencies.

Need to compare Carweek with other indicators of driving in last week:

* MinsCar
* Diary1car, Diary2car, Diary3car, Diary4car, Diary5car, Diary6car, Diary7car
* Car1min, Car2min, Car3min, Car4min, Car5min, Car6min, Car7min


## Carweek 'refused' check

Are any Carweek entries entered as 'refused', other strings or NA?

```{r Carweek entries}
## Check 'refused' responses - should just be 0 or 1
distinct(select(pp, Carweek))
```
There are no 'refused' entries in Carweek. All coded as 0 or 1.


## Carweek=0 and Diary?car=1 check.

Where Carweek=0, are there any driving diary entries that are 1 (Diary?car=1).

```{r NO Carweek but YES Diary?car}
## Check dimensions of Carweek=0 & Diary?car=1
ppv %>%
  filter(Carweek==0 | is.na(Carweek), Diary1car==1 | Diary2car==1 | Diary3car==1 | Diary4car==1 | Diary5car==1 | Diary6car==1 | Diary7car==1) %>%
  select(ID, Carweek, Diary1car, Diary2car, Diary3car, Diary4car, Diary5car, Diary6car, Diary7car) %>%
  View()
```
There are 138 records where Carweek=0 or is NA that have an associated driving diary entry.


Add 'noCarweek_yesDiary' variable flagging these.
```{r Add noCarweek_yesDiary variable}
## Add flagging variable
ppv <- ppv %>%
  mutate(noCarweek_yesDiary = ifelse((Carweek==0 | is.na(Carweek)) & (Diary1car==1 | Diary2car==1 | Diary3car==1 | Diary4car==1 | Diary5car==1 | Diary6car==1 | Diary7car==1), 1, NA))

## Check new variable
ppv %>%
  filter(Carweek==0, Diary1car==1 | Diary2car==1 | Diary3car==1 | Diary4car==1 | Diary5car==1 | Diary6car==1 | Diary7car==1) %>%
  select(ID, noCarweek_yesDiary, Carweek, Diary1car) %>%
  View()
```

## Carweek=1 and Diary?car=0 check.

Where Carweek=1 are there any driving diary entries that are zero (Diary?car=0).
```{r YES Carweek but NO Diary?car}
ppv %>%
  filter(Carweek==1, (Diary1car==0 | is.na(Diary1car)) & (Diary2car==0 | is.na(Diary2car)) & (Diary3car==0 | is.na(Diary3car)) & (Diary4car==0 | is.na(Diary4car)) & (Diary5car==0 | is.na(Diary5car)) & (Diary6car==0 | is.na(Diary6car)) & (Diary7car==0 | is.na(Diary7car))) %>%
  select(ID, Carweek, Diary1car, Diary2car, Diary3car, Diary4car, Diary5car, Diary6car, Diary7car) %>%
  View()
```
There are zero records where Carweek=1 and no driving diary entry.


## Carweek and MinsCar variable check.

What variables have been entered in MinsCar (Car minutes during last week). This should just be numeric but are there NAs or letters?

```{r MinsCar entries}
## Check for NAs
ppv %>%
  filter(is.na(MinsCar)) %>%
  select(ID, Carweek, MinsCar, Car1min:Car7min)
## No NAs

## Check for letters in MinsCar responses - should just be numbers
## ref http://stackoverflow.com/a/21629102 to get character vector from tbl_df
mc <- collect(select(pp,MinsCar))[[1]]

pattern <- "[A-z]"

mc[str_detect(mc, pattern)]

rm(mc, pattern)
```
No NAs or letter in MinsCar.


## Carweek=0 and MinsCar>0 check.

Are there instances of Carweek=0 & MinsCar>0?
```{r NO Carweek but YES MinsCar}
pp %>%
  filter(Carweek==0 | is.na(Carweek), MinsCar>0) %>%
  dim()
```
There are zero records of Carweek=0 with MinsCar>0.

## Carweek=1 and MinsCar=0 or NA check.

Are there instances of Carweek=1 with MinsCar=0 or NA?
```{r YES Carweek but NO MinsCar}
pp %>%
  filter(Carweek==1, MinsCar==0 | is.na(MinsCar)) %>%
  select(ID, Carweek, MinsCar) %>%
  View()
```
There are 36 records of Carweek=1 and MinsCar=0.

Add 'yesCarweek_noMinsCar' variable flagging these.
```{r Add yesCarweek_noMinsCar variable}
ppv <- ppv %>%
  mutate(yesCarweek_noMinsCar = ifelse(Carweek==1 & (MinsCar==0 | is.na(MinsCar)), 1, NA))

ppv %>%
  select(ID, yesCarweek_noMinsCar, Carweek, MinsCar) %>%
  filter(yesCarweek_noMinsCar==1) %>%
  View()

## Save data
#save.image("~/Documents/Personal/Work/DataScience/PeopleAndPlaces/Data/pp.data.RData")

```

## Carweek and Car?min variable check.

Check that where Carweek=0 doesn't have any Car?min entries.
```{r Carweek and Car?min NA check}
## Check Carweek==0 but Car?min... are not NA
ppv %>%
  select(ID, Carweek, Car1min:Car7min) %>%
  filter(Carweek==0, !is.na(Car1min) | !is.na(Car2min) | !is.na(Car3min) | !is.na(Car4min) | !is.na(Car5min) | !is.na(Car6min) | !is.na(Car7min)) %>%
  View()
```
No Car?min entries with Carweek=0.


## Carweek=1 and Car?min=0 or NA check.
Where Carweek=1, are there NA entries to driving minutes per day (Car?min) variables?
```{r YES Carweek but NO Car?min}
ppv %>%
  select(ID, Carweek, Car1min:Car7min) %>%
  filter(Carweek==1, (is.na(Car1min) | Car1min==0) & (is.na(Car2min) | Car2min==0) & (is.na(Car3min) | Car3min==0) & (is.na(Car4min) | Car4min==0) & (is.na(Car5min) | Car5min==0) & (is.na(Car6min) | Car6min==0) & (is.na(Car7min) | Car7min==0)) %>%
  View()

## 36 records
```
There are 36 instances of Carweek=1 with NA or 0 values for driving minutes.


Add 'yesCarweek_noCarmins' variable flagging these.
```{r Add yesCarweek_noCarmins variable}
## Add flagging variable
ppv <- ppv %>%
  mutate(yesCarweek_noCarmins = ifelse(Carweek==1 & (is.na(Car1min) | Car1min==0) & (is.na(Car2min) | Car2min==0) & (is.na(Car3min) | Car3min==0) & (is.na(Car4min) | Car4min==0) & (is.na(Car5min) | Car5min==0) & (is.na(Car6min) | Car6min==0) & (is.na(Car7min) | Car7min==0), 1, NA))

ppv %>%
  select(ID, yesCarweek_noCarmins, Carweek, Car1min) %>%
  filter(yesCarweek_noCarmins == 1) %>%
  View()
```

? are yesCarweek_noCarmins the same records as yesCarweek_noMinsCar records?
```{r Join by ID}
ppv.MinsCar <- ppv %>%
  select(ID, yesCarweek_noMinsCar) %>%
  filter(yesCarweek_noMinsCar==1) %>%
  mutate(type = "MinsCar")

ppv.Carmin <- ppv %>%
  select(ID, yesCarweek_noCarmins) %>%
  filter(yesCarweek_noCarmins==1) %>%
  mutate(type = "Carmins")

join <- inner_join(ppv.Carmin, ppv.MinsCar, by="ID")
  
View(join)

## Yes they are the same

#rm(join, ppv.Carmin, ppv.MinsCar)
```

## Combine yesCarweek_noMinsCar and yesCarweek_noCarmins

As the combination of Carweek=1 and Car?mins=0 or NA, and Carweek=1 and MinsCar=0 or NA, yield the same records, they can be combined together into the same flagging variable.

```{r Combine yesCarweek_noMinsCar and yesCarweek_noCarmins}
## Remove yesCarweek_noMinsCar and rename yesCarweek_noCarmins variables
ppv <- ppv %>%
  select(-yesCarweek_noMinsCar) %>%
  rename(yesCarweek_noMins = yesCarweek_noCarmins)

ppv %>%
  filter(yesCarweek_noMins==1) %>%
  select(ID, Cycleweek, yesCarweek_noMins, MinsCar, Car1min) %>%
  View()
```

## Carweek=0 and Car?min=1 check.
```{r NO Carweek and YES Car?min}
# Check Carweek==0 but Car?min... are not zero
ppv %>%
  select(ID, Carweek, Car1min:Car7min) %>%
  filter((Carweek==0 | is.na(Carweek)), Car1min>0 & Car2min>0 & Car3min>0 & Car4min>0 & Car5min>0 & Car6min>0 & Car7min>0) %>%
  dim()
```




















